---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(coronavirus)
library(tidyverse)
library(incidence)
library(ggthemes)
library(plotly)
# filter global data for region of interest

```

```{r}
coronavirus_dataset <- read_csv("https://github.com/RamiKrispin/coronavirus-csv/raw/master/coronavirus_dataset.csv")

```



```{r}
df_usa <-
  coronavirus %>%
  filter(Country.Region == "US",
         !Province.State %in% c("Diamond Princess", "Grand Princess"))

df_usa %>%
filter(type == "confirmed") %>%
group_by(date) %>%
summarise(incident_cases = sum(cases)) %>%
mutate(cumulative_cases = cumsum(incident_cases)) %>%
ggplot() +
geom_area(aes(x = date, y = cumulative_cases), stat = "identity", fill = "sienna")+
geom_bar(aes(x = date, y = incident_cases), stat = "identity", fill = "blue") +
labs(y = "", title = "COVID-19 Cases", x = "") +
theme_pander() 

```

```{r}
df_usa %>%
  select(date, type, everything(), -Province.State) %>% 
arrange(date) %>%
group_by(date, type) %>%
summarise(cases = sum(cases)) %>%
pivot_wider(date, names_from = "type", values_from = "cases") %>%
ungroup() %>%
mutate(
  active = confirmed - death,# - recovered,
confirmed_cum = cumsum(confirmed),
death_cum = cumsum(death),
# recovered_cum = cumsum(recovered),
active_cum = cumsum(active)) %>%
select(date, 
       active_cum, 
       # recovered_cum, 
       death_cum) %>%
pivot_longer(-date, names_to = "type", values_to = "cases") %>%
ggplot(aes(x = date, y = cases, colour = type)) +
geom_line(size = 1.5) +
scale_color_hue() +
labs(y = "Cumulative Cases", title = "Cumulative Active and Death Cases") +
  theme_pander()
```

Let's filter our full data down to confirmed cases and sum them according to date. 

```{r}
df_confirmed <- 
  df_usa %>%
  filter(type == "confirmed") %>%
  group_by(date) %>%
  summarise(incident_cases = sum(cases))
```

Next we create an object of class `incidence` using the `incidence()` function.  Suppose we wished to view the incidence by week, or every 7 days. The `incidence()` function wants a vector of observation so we `pull()` the `incident_cases` column from our `df_confirmed` tibble.

```{r}

dates_cases_tbl <- 
  rep(df_confirmed$date, df_confirmed$incident_cases) %>% 
  enframe(name = "day_num") %>% 
  rename(date_of_obs = value) %>% 
  group_by(date_of_obs) %>% 
  mutate(total_cases_on_date = length(day_num))

dates_cases_tlb
```
We went through reorganizing the data in a `date_cases_tbl` that we could `filter()` by the `date_of_obs` column inside our call to the `incidence()` function. I prefer this because we can see the dates more clearly and it will help with Shiny later when we want to sort by date, not by row index.

```{r}
days_interval <- 2
start_date <- lubridate::ymd("2020-02-20")

incidence(dates_cases_vector %>% filter(date_of_obs > start_date) %>% pull(date_of_obs), interval = days_interval) %>% 
  plot(color = "blue") +
  labs(y = "", title = str_glue("Incidence Every {days_interval} Days"), caption = "source: rami's package") +
  xlim(start_date, lubridate::ymd("2020-03-30")) +
  theme_pander()
```

```{r}
early_mod_start_date <- lubridate::ymd("2020-03-02")
early_mod_end_date <- lubridate::ymd("2020-03-11")

jkr_early_fit <-
  fit(incidence(dates_cases_vector %>% filter(between(date_of_obs, early_mod_start_date, early_mod_end_date)) %>% pull(date_of_obs)), split = NULL)

```

Let's pass our results to `ggplot()` and look at the predicted values, along with lower and upper confidence intervals.

```{r}
jkr_early_fit$info$pred %>% 
  select(-dates.x) %>% 
  pivot_longer(-dates, names_to = "type", values_to = "result") %>% 
  ggplot(aes(x = dates, y = result, color = type, linetype = type)) + 
  geom_line() 
```

We can add data from our `df_confirmed` table so we can compare the predicted levels to actual levels.

```{r}
jkr_early_fit$info$pred %>% 
  select(-dates.x) %>% 
  pivot_longer(-dates, names_to = "type", values_to = "result") %>% 
  ggplot(aes(x = dates, y = result, color = type, linetype = type)) + 
  geom_line() +
  geom_bar(data = df_confirmed %>% filter(date >= early_mod_start_date, date <= early_mod_end_date), 
           aes(x = date, weight = incident_cases),
           alpha = .7,
           fill = "lightgrey", 
           inherit.aes = FALSE) +
labs(y = "", x = "", title = "Confirmed Cases v. Early Model Prediction") +
scale_x_date(labels = scales::date_format("%d %b %Y")) +
ggthemes::theme_pander()
```

Since the model results are stored in a standard `model` object, we can use `tidy()` from the `broom` package to view our coefficients.

```{r}

jkr_early_fit$model %>% 
  broom::tidy()
```

Let's pull out the coefficients for each of our terms, the date and the intercept and save them as separate variables. 

```{r}
early_model_date_coefficient <- 
  jkr_early_fit$model %>% 
  broom::tidy() %>% 
  filter(term == "dates.x") %>% 
  pull(estimate)

early_model_intercept_coefficient <- 
  jkr_early_fit$model %>% 
  broom::tidy() %>% 
  filter(term == "(Intercept)") %>% 
  pull(estimate)


```

Now let's use those coefficients to get predictions from our model. First we determine many periods we want to forecast, and use that number to build a sequence of dates.

```{r}
n_forecasts <- 
  as.numeric(difftime(max(df_confirmed$date), jkr_early_fit$origin, units = "days")) + 1

forecast_dates <- seq.Date(from = jkr_early_fit$origin, length.out = n_forecasts, by = "days")

early_model_predictions <- exp(early_model_intercept_coefficient + early_model_date_coefficient * (0:(n_forecasts - 1)))

model_pred_df <- tibble(date = forecast_dates, prediction = early_model_predictions)
```


```{r}
ggplot() +
geom_bar(data = df_confirmed %>% filter(date >= jkr_early_fit$origin),
aes(x = date, y = incident_cases), stat = "Identity", fill = "darkgray") +
geom_line(data = model_pred_df, aes(x = date, y = prediction), size = 1L, colour = "blue") +
labs(y = "Incidence", title = "Actual and Predicted Incidence Rates") +
ggthemes::theme_pander()
```
Now let's fit a model to our entire data set. We do this by setting a start date to the same as our early model with `early_mod_start_date` but extending our end date to today, with `Sys.Date()`. 

```{r}

jkr_current_fit <-
  fit(incidence(dates_cases_vector %>% filter(between(date_of_obs, early_mod_start_date, Sys.Date())) %>% pull(date_of_obs)), split = NULL)

```


```{r}

jkr_current_fit$info$pred %>% 
  select(-dates.x) %>% 
  pivot_longer(-dates, names_to = "type", values_to = "result") %>% 
  ggplot(aes(x = dates, y = result, color = type, linetype = type)) + 
  geom_line() +
  geom_bar(data = df_confirmed %>% filter(date >= start_date), 
           aes(x = date, weight = incident_cases),
           alpha = .7,
           fill = "lightgrey", 
           inherit.aes = FALSE) +
labs(y = "", x = "", title = "Confirmed Cases v. Model Prediction") +
scale_x_date(labels = scales::date_format("%d %b %Y")) +
ggthemes::theme_pander()
```


Let's use `tidy()` from the `broom` package to glance at our coefficients.

```{r}

jkr_current_fit$model %>% 
  broom::tidy()
```

Let's pull out the coefficients for each of our terms, the date and the intercept.

```{r}
jkr_current_model_date_coefficient <- 
  jkr_current_fit$model %>% 
  broom::tidy() %>% 
  filter(term == "dates.x") %>% 
  pull(estimate)

jkr_current_model_intercept_coefficient <- 
  jkr_current_fit$model %>% 
  broom::tidy() %>% 
  filter(term == "(Intercept)") %>% 
  pull(estimate)


```

We will extend our forecast out to 4 days in the future, instead of just 1 day as we did previously. 

```{r}
n_forecasts <- 
  as.numeric(difftime(max(df_confirmed$date), jkr_current_fit$origin, units = "days")) + 4

forecast_dates <- seq.Date(from = jkr_current_fit$origin, length.out = n_forecasts, by = "days")

current_model_predictions <- exp(jkr_current_model_intercept_coefficient + jkr_current_model_date_coefficient * (0:(n_forecasts - 1)))

early_model_predictions_extended <- exp(early_model_intercept_coefficient + early_model_date_coefficient * (0:(n_forecasts - 1)))

model_pred_tbl <- 
  tibble(date = forecast_dates,
         early_model = early_model_predictions_extended,
         current_model = current_model_predictions)

plot_data_tbl <- left_join(model_pred_tbl, df_confirmed)

```



```{r}
ggplot() +
  geom_bar(
    data = plot_data_tbl,
    aes(x = date, y = incident_cases),
    stat = "Identity",
    fill = "darkgray"
  ) +
  geom_line(data = plot_data_tbl,
            aes(x = date, y = early_model, color = "Early Model"),
            size = 1L) +
  geom_line(
    data = plot_data_tbl,
    aes(x = date, y = current_model, color = "Current Model"),
    size = 1L
  ) +
  labs(y = "Incidence", title = "Actual and Predicted Incidence Rates", color = 'Model') +
  scale_color_manual(values = c('Early Model' = 'darkblue',
                                'Current Model' = 'orange')) +
  xlim(jkr_current_fit$origin, lubridate::ymd("2020-03-30")) +
  ggthemes::theme_pander()
```
Finally, let's take a look at how our early model and current model compare in their predictions for the next few days.

```{r}
plot_data_tbl %>% 
  tail(7)
```

