---
title: "Factor Notebook"
output: html_notebook
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, messaage = FALSE, warning = FALSE, comment = NA)
library(tidyverse)
library(tidyquant)
library(tibbletime)
library(broom)
```

# Get the data

```{r}
combined_ts <- suppressMessages(read_csv("combined.ts.csv" )) %>% 
  na.omit()
```


# Rolling Regression
 
Create the rolling function
```{r}
rolling_lm <- rollify(.f = function(Asset, Growth, Value, Momentum, Size, ERP) {

  lm(Asset ~ Growth + Value + Momentum + Size + ERP)

},

window = 36,

unlist = FALSE)
```

Use the ` %>% ` operator to apply that rolling function to the data and look at the results. 

```{r}
combined_ts %>% 
   mutate(rolling_model = 
           rolling_lm(Fund, 
                      US.Pure.Growth, 
                      US.Pure.Value, 
                      US.Pure.Momentum,
                      US.Pure.Size,
                      RMRF)) %>% 
  # Why slice of 35 rows? It's a rolling model of size 36, so the first 35 are NA
  slice(-1:-35) %>%
  # look at just the datestamp and model results column
  select(datestamp, rolling_model)  %>% 
  head()
```

Look at the new column called rolling_model.
It has the results of our rolling model, now we need to extract the results for that column. We do that with `tidy()` function from `broom`. But, remember, we are applying it iteratively to objects in a column called `rolling_model`, so we need to use `map()` to apply the `tidy()` function to that column. Then we `unnest()` the results.

```{r}

model_results <- 
  combined_ts %>% 
  mutate(rolling_model = 
           rolling_lm(Fund, 
                      US.Pure.Growth, 
                      US.Pure.Value, 
                      US.Pure.Momentum,
                      US.Pure.Size,
                      RMRF)) %>% 
  mutate(tidied = map(rolling_model, 
                      tidy, 
                      conf.int = T)) %>% 
  unnest(tidied) %>% 
  # Slice off the NAs again
  slice(-1:-35) %>% 
  select(datestamp, term, estimate, conf.low, conf.high) %>% 
  filter(term != "(Intercept)") %>% 
  rename(beta = estimate, factor = term) %>% 
  group_by(factor)

model_results %>% 
  head()
```

# Visualize results

```{r}
model_results %>% 
  ggplot(aes(x = datestamp, 
             y = beta, 
             group = factor, 
             color = factor)) +
  geom_line()
  
```

