---
title: "jobs-report-vis-from-sql"
resource_files: config.yml
output:
  html_document:
    df_print: paged
rmd_output_metadata:
  rsc_email_suppress_scheduled: true
---

```{r setup, include = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)
library(tidyverse)
library(tidyquant)
library(highcharter)
library(openxlsx)

library(config)
library(dbplyr)
library(DBI)
library(odbc)
library(htmltools)

dw <- config::get("mssql")

con <- DBI::dbConnect(odbc::odbc(),
                      Driver = dw$Driver,
                      Server = dw$Server,
                      #Host = dw$Host,
                      UID    = dw$UID,
                      PWD    = dw$PWD,
                      Port   = dw$Port,
                      Database = dw$Database
                      #Schema = dw$Schema
                      )

source("recession-data/recession-geom.R")
```

### Import data from sql

```{r}
bls_labor_for_data <-
  tbl(con, "bls_labor_for_data") %>% 
  collect() %>% 
  mutate(date = ymd(date),
         value = as.numeric(value))

fred_empl_data <-
  tbl(con, "fred_empl_data") %>% 
  collect()
```


### Chop up the Fred data into sectors and hours/wages objects

```{r}

labor_sector_data <- 
  fred_empl_data  %>% 
  filter(date >= "1990-01-01") %>%
  select(1, 9:22) %>% 
  gather(sector, employees, -date) %>% 
  group_by(sector) %>% 
  mutate(monthly_change = employees - lag(employees, 1),
         yearly_change = employees - lag(employees, 12)) %>%
  na.omit()

hours_wages_data <-
  fred_empl_data%>%
  select(date, `Avg Hourly Earnings`, `Avg Weekly Hours`)  %>% 
  na.omit()

```

### Visualize Total Employment Since 2009 with ggplot

```{r}
labor_sector_data %>% 
  filter(sector == "Nonfarm Employment") %>% 
  filter(date >= "2009-01-01") %>% 
   mutate(col_blue = 
           if_else(monthly_change > 0, 
                  monthly_change, as.numeric(NA)),
         col_red = 
           if_else(monthly_change < 0, 
                  monthly_change, as.numeric(NA))) %>%
  ggplot(aes(x = date)) +
  geom_col(aes(y = col_red),
               alpha = .85, 
               fill = "pink", 
               color = "pink") +
  geom_col(aes(y = col_blue),
               alpha = .85, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  ylab("Monthly Change (thousands)") +
  labs(title = "Monthly Private Employment Change",
       subtitle = "total empl, since 2009",
       caption = "based on work by @lenkiefer") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption=element_text(hjust=0))
```

### Visualize Total Employment Since 2009 with highcharter

```{r}
labor_data_hc <- 
  labor_sector_data %>% 
  filter(sector == "Nonfarm Employment") %>% 
  filter(date >= "2009-01-01") %>%
   mutate(coloract = if_else(monthly_change > 0, "#6495ed", "#ffe6ea"))

highchart() %>%
  hc_title(text = "Monthly Employment Change") %>%
  hc_add_series(labor_data_hc,  
                "column", 
                pointWidth = 5,
                hcaes(x = date,
                      y = monthly_change,
                      color = coloract),
                name = "monthly empl change") %>% 
  hc_xAxis(type = "datetime") %>% 
  hc_exporting(enabled = TRUE)

```


### Visualize Sector by Sector Change Last Month with ggplot


```{r}
labor_sector_data  %>% 
  filter(date == (last(date))) %>% 
   mutate(col_blue = 
           if_else(monthly_change > 0, 
                  monthly_change, as.numeric(NA)),
         col_red = 
           if_else(monthly_change < 0, 
                  monthly_change, as.numeric(NA))) %>%
  ggplot(aes(x = reorder(sector, monthly_change))) +
  geom_col(aes(y = col_red),
               alpha = .85, 
               fill = "pink", 
               color = "pink") +
  geom_col(aes(y = col_blue),
               alpha = .85, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  labs(title = paste((lubridate::month(last(labor_sector_data$date), 
                                       label = TRUE, abbr = FALSE)), 
                     "Employment Change", 
                     sep = " "),
       subtitle = "by sector, and total", 
       x = "sector", 
       y = "Monthly Change (thousands)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

### Visualize Sector by Sector Change Last Month with highcharter

```{r}
  
labor_sector_most_recent_month <- 
  labor_sector_data  %>% 
  filter(date == (last(date))) %>%
  arrange(monthly_change) %>% 
  mutate(coloract = if_else(monthly_change > 0, "#6495ed", "#ffe6ea"))



highchart() %>%
  hc_title(text = paste(lubridate::month(last(labor_sector_data$date),
                                 label = TRUE, 
                                 abbr = FALSE), 
                        "Employment Change", 
                        sep = " ")) %>%
  hc_add_series(labor_sector_most_recent_month,  
                "column", 
                pointWidth = 20,
                hcaes(x = sector,
                      y = monthly_change,
                      color = coloract),
                showInLegend = FALSE) %>% 
  hc_xAxis(categories = labor_sector_most_recent_month$sector) %>%
  hc_xAxis(title = list(text = "Sector")) %>% 
  hc_yAxis(title = list(text = "Monthly Change (thousands)"))


```


### Visualize Sector by Sector Data over time with ggplot

```{r, fig.height=10, fig.width=12}

labor_sector_data %>% 
  filter(sector != "Nonfarm Employment") %>% 
  filter(date >= (last(date) - months(24))) %>% 
   mutate(col_blue = 
           if_else(monthly_change > 0, 
                  monthly_change, as.numeric(NA)),
         col_red = 
           if_else(monthly_change < 0, 
                  monthly_change, as.numeric(NA))) %>%
  ggplot(aes(x = date)) +
  geom_col(aes(y = col_red),
               alpha = .85, 
               fill = "pink", 
               color = "pink") +
  geom_col(aes(y = col_blue),
               alpha = .85, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  ylab("Monthly Change (thousands)") +
  labs(title = "Monthly Private Employment Change",
       subtitle = "by sector, 24 months") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
    strip.background = element_blank(),
    strip.placement = "inside",
    strip.text = element_text(size=15),
    panel.spacing = unit(0.2, "lines") ,
    panel.background=element_rect(fill="white")) +
  facet_wrap(~sector, shrink = FALSE)
```


### Visualize Sector by Sector Data over time with highcharter
```{r}
labor_sector_data_spread_hc <- 
  labor_sector_data %>% 
  filter(sector != "Nonfarm Employment") %>% 
  filter(date >= (last(date) - months(24))) %>% 
  select(date, sector, monthly_change) %>% 
  spread(sector, monthly_change)

map(names(labor_sector_data_spread_hc[2:11]), function(x){

labor_data_hc_fun <- 
  labor_sector_data %>% 
  filter(sector == x) %>% 
  filter(date >= (last(date) - months(24))) %>% 
   mutate(coloract = if_else(monthly_change > 0, "#6495ed", "#ffe6ea"))

highchart() %>%
  hc_title(text = paste(x, "Monthly Employment Change", sep = " ")) %>%
  hc_add_series(labor_data_hc_fun,  
                "column", 
                pointWidth = 3,
                hcaes(x = date,
                      y = monthly_change,
                      color = coloract),
                name = "monthly empl change") %>% 
  hc_xAxis(type = "datetime") %>% 
  hc_legend(enabled = FALSE) %>% 
  hc_exporting(enabled = TRUE)
}) %>% 
  hw_grid(rowheight = 300, ncol = 3) %>% 
  browsable()
```



### Visualize Wage and Hours Changes

```{r}
hours_wages_data %>% 
  gather(stat, obs, -date) %>% 
  group_by(stat) %>% 
  mutate(monthly_change = obs - lag(obs, 1)) %>% 
  filter(date >= last(date) - months(12)) %>%
   mutate(col_blue = 
           if_else(monthly_change > 0, 
                  monthly_change, as.numeric(NA)),
         col_red = 
           if_else(monthly_change < 0, 
                  monthly_change, as.numeric(NA))) %>%
  ggplot(aes(x = date)) +
  geom_col(aes(y = col_red),
               alpha = .85, 
               fill = "pink", 
               color = "pink") +
  geom_col(aes(y = col_blue),
               alpha = .85, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  ylab("Monthly Change") +
  labs(title = "Wages and Hours",
       subtitle = "last 12 months") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
    strip.background = element_blank(),
    strip.placement = "inside",
    strip.text = element_text(size=15),
    panel.spacing = unit(0.2, "lines") ,
    panel.background=element_rect(fill="white")) +
  facet_wrap(~stat, shrink = FALSE)
```

### Wrangle the macro economic data

```{r}
fred_empl_data_wrangled <- 
  fred_empl_data  %>%
  select(date, 
         `Nat Rate Unemployment`, 
         `Unemployment Rate`,
         `Nonfarm Employment`,
         `Labor Force Part Rate`,
         `Employment-to-Pop Ratio`,
         `Nonfarm Employment`) %>% 
  mutate(`Nat Rate Unemployment` = na.locf(`Nat Rate Unemployment`, na.rm = F)) %>%
  mutate(`Unemployment Gap` = `Unemployment Rate` - `Nat Rate Unemployment`,
         `Employment Monthly Change` = `Nonfarm Employment` - lag(`Nonfarm Employment`))

```

### Vis Showing Above or Below Trend

```{r}
fred_empl_data_wrangled %>% 
  mutate(ribbon_pink = 
           if_else(`Unemployment Rate` > `Nat Rate Unemployment`, 
                  `Unemployment Rate`, `Nat Rate Unemployment`),
         ribbon_blue = 
           if_else(`Unemployment Rate` < `Nat Rate Unemployment`, 
                  `Unemployment Rate`, `Nat Rate Unemployment`)) %>% 
  ggplot(aes(x = date, y = `Unemployment Rate`)) +
  geom_line(color = "black") +
  geom_line(linetype = 2, aes(y = `Nat Rate Unemployment`))  +
  recession_shade +
  geom_ribbon(aes(ymin = `Nat Rate Unemployment`, 
                  ymax = ribbon_pink), 
              fill = "pink", 
              alpha = 0.5)+
  geom_ribbon(aes(ymin = `Nat Rate Unemployment`,
                  ymax = ribbon_blue), 
              fill = "cornflowerblue", 
              alpha = 0.5)  +
  xlim(ymd("1980-01-01"), ymd("2018-03-01")) +
  labs(title = "Unemployment v. Nat Rate Unemploymnet",
       caption = "data source: fred",
       y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.caption  = element_text(hjust = 0)) 
```

### Grab the ADP and BLS Data from the Fred object

```{r}
adp_fred <-
  fred_empl_data  %>%
  select(date, ADP = `ADP Estimate`, BLS = `Nonfarm Employment`)  %>% 
    gather(report, month, -date) %>% 
    group_by(report) %>% 
    mutate(monthly_change = month - lag(month, 1))
```

```{r}
adp_fred %>% 
  filter(date >= "2017-01-01") %>%
  ggplot(aes(x = date, y = monthly_change, fill = report)) + 
  geom_col(position = "dodge") + 
  scale_fill_manual(values = c("pink","cornflowerblue")) +
  labs(title = "ADP Estimate v. BLS Report",
       caption = "data source: fred",
       y = "monthly change (thousands)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.caption  = element_text(hjust = 0)) +
  scale_x_date(breaks = scales::pretty_breaks(n = 16))
```


### Vis from BLS direct download

```{r}
bls_labor_for_data %>% 
  group_by(dem) %>%
  filter(date >= "1980-01-01") %>% 
  ggplot(aes(x = date, y = value, color = dem, label = dem)) +
  recession_shade +
  geom_line()  +
  theme_minimal() +
  scale_x_date(date_breaks = "2 years", 
               date_labels = "%Y") +
  xlim(ymd("1980-01-01"), Sys.Date()) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10), 
                     labels = function(x) paste0(x, "%"), 
                     sec.axis = dup_axis()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption=element_text(hjust=0)) +
  ylab("") +
  xlab("") +
  ggtitle("Labor Force Participation Rate: Prime Working Age (25-54)", 
          subtitle = "seasonally adjusted") +
  labs(caption = "based on work by the inestimable @lenkiefer") 
```



```{r}
library(timetk)
bls_labor_for_data_xts <- 
  bls_labor_for_data %>% 
  group_by(dem) %>%
  filter(date >= "1980-01-01") %>% 
  spread(dem, value) %>% 
  tk_xts(date_var = date)


highchart(type = 'stock')%>%
  hc_add_series(bls_labor_for_data_xts$All, name = "All") %>%
  hc_add_series(bls_labor_for_data_xts$Men, name = "Men") %>%
  hc_add_series(bls_labor_for_data_xts$Women, name = "Women") %>%
  hc_title(text = "Labor Force Participation") %>% 
      hc_xAxis(type = 'datetime',
                 plotBands = list(
                   list(
                     label = list(text = "Bush Senior"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('1990-07-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('1991-03-01', tz = 'UTC'))
                     ),
                   list(
                     label = list(text = "Internet Bubble"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('2001-03-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('2001-11-01', tz = 'UTC'))
                     ),
                   list(
                     label = list(text = "Financial Crisis"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('2007-12-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('2009-06-01', tz = 'UTC'))
                     ),
                   list(
                     label = list(text = "Volker Rate Rise"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('1981-07-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('1982-11-01', tz = 'UTC'))
                     )
                   )
               ) %>% 
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_exporting(enabled = TRUE)

```

```{r, echo = FALSE}
labor_for_part_fname <- 
  paste("labor_for_part_",Sys.Date(), ".xlsx", sep = "")

write.xlsx(bls_labor_for_data_xts, file = labor_for_part_fname)

rmarkdown::output_metadata$set(rsc_output_files = list(labor_for_part_fname))
```

[Link to labor force part data Excel](`r labor_for_part_fname`)


