---
title: "Jobs Report"
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
rmd_output_metadata:
  rsc_email_suppress_scheduled: true
---

```{r setup, include = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
library(tidyverse)
library(tidyquant)
```


### Import data from Fred

```{r}
fred_codes <- c(
            "CIVPART", 
            "EMRATIO",
            "NROU",  
            "UNRATE", 
            "CES0500000003",
            "AWHAETP",
            "NPPTTL",
            "PAYEMS", 
            "USCONS",
            "USTRADE",
            "USPBS",
            "MANEMP",
            "USFIRE",
            "USMINE",
            "USEHS",
            "USWTRADE",
            "CES4349300001",
            "USINFO",
            "USLAH",
            "USGOVT",
            "USSERV")
            
column_names <- c(
            "date",
            "Labor Force Part Rate",
            "Employment-to-Pop Ratio",
            "Nat Rate Unemployment",
            "Unemployment Rate",
            "Avg Hourly Earnings",
            "Avg Weekly Hours",
            "ADP Estimate",
            "Nonfarm Employment",
            "Construction",
            "Retail/Trade",
            "Prof/Bus Serv",
            "Manufact",
            "Financial",
            "Mining",
            "Health Care",
            "Wholesale Trade",
            "Transportation",
            "Info Sys",
            "Leisure",
            "Gov",
            "Other Services")


# download data via FRED 
fred_empl_data <- 
  tq_get(fred_codes,                         
         get = "economic.data",             
         from = "1980-01-01") %>% 
  spread(symbol, price) %>% 
  dplyr::select(date, fred_codes) %>% 
  `colnames<-`(column_names)
```



### Chop up the Fred data into sectors and hours/wages objects

```{r}

labor_sector_data <- 
  fred_empl_data  %>% 
  filter(date >= "1990-01-01") %>%
  select(1, 9:22) %>% 
  gather(sector, employees, -date) %>% 
  group_by(sector) %>% 
  mutate(monthly_change = employees - lag(employees, 1),
         yearly_change = employees - lag(employees, 12)) %>%
  na.omit()

hours_wages_data <-
  fred_empl_data%>%
  select(date, `Avg Hourly Earnings`, `Avg Weekly Hours`)  %>% 
  na.omit()

```

### Visualize Total Employment Since 2009

```{r}
labor_sector_data %>% 
  filter(sector == "Nonfarm Employment") %>% 
  filter(date >= "2009-01-01") %>% 
   mutate(col_blue = 
           if_else(monthly_change > 0, 
                  monthly_change, as.numeric(NA)),
         col_red = 
           if_else(monthly_change < 0, 
                  monthly_change, as.numeric(NA))) %>%
  ggplot(aes(x = date)) +
  geom_col(aes(y = col_red),
               alpha = .85, 
               fill = "blue", 
               color = "blue") +
  geom_col(aes(y = col_blue),
               alpha = .85, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  ylab("Monthly Change (thousands)") +
  labs(title = "Monthly Private Employment Change",
       subtitle = "total empl, since 2009",
       caption = "based on work by @lenkiefer") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption=element_text(hjust=0))
```

### Visualize Sector by Sector Change Last Month


```{r}
labor_sector_data  %>% 
  filter(date == (last(date))) %>% 
   mutate(col_blue = 
           if_else(monthly_change > 0, 
                  monthly_change, as.numeric(NA)),
         col_red = 
           if_else(monthly_change < 0, 
                  monthly_change, as.numeric(NA))) %>%
  ggplot(aes(x = reorder(sector, monthly_change))) +
  geom_col(aes(y = col_red),
               alpha = .85, 
               fill = "blue", 
               color = "blue") +
  geom_col(aes(y = col_blue),
               alpha = .85, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  labs(title = paste((lubridate::month(last(labor_sector_data$date), 
                                       label = TRUE, abbr = FALSE)), 
                     "Employment Change", 
                     sep=" "),
       subtitle = "by sector, and total", x = "sector", y= "Monthly Change (thousands)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

### Visualize Sector by Sector Data over time

```{r, fig.height=10, fig.width=12}

labor_sector_data %>% 
  filter(sector != "Nonfarm Employment") %>% 
  filter(date >= (last(date) - months(24))) %>% 
   mutate(col_blue = 
           if_else(monthly_change > 0, 
                  monthly_change, as.numeric(NA)),
         col_red = 
           if_else(monthly_change < 0, 
                  monthly_change, as.numeric(NA))) %>%
  ggplot(aes(x = date)) +
  geom_col(aes(y = col_red),
               alpha = .85, 
               fill = "blue", 
               color = "blue") +
  geom_col(aes(y = col_blue),
               alpha = .85, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  ylab("Monthly Change (thousands)") +
  labs(title = "Monthly Private Employment Change",
       subtitle = "by sector, 24 months") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
    strip.background = element_blank(),
    strip.placement = "inside",
    strip.text = element_text(size=15),
    panel.spacing = unit(0.2, "lines") ,
    panel.background=element_rect(fill="white")) +
  facet_wrap(~sector, shrink = FALSE)
```

### Visualize Wage and Hours Changes

```{r}
hours_wages_data %>% 
  gather(stat, obs, -date) %>% 
  group_by(stat) %>% 
  mutate(monthly_change = obs - lag(obs, 1)) %>% 
  filter(date >= last(date) - months(12)) %>%
   mutate(col_blue = 
           if_else(monthly_change > 0, 
                  monthly_change, as.numeric(NA)),
         col_red = 
           if_else(monthly_change < 0, 
                  monthly_change, as.numeric(NA))) %>%
  ggplot(aes(x = date)) +
  geom_col(aes(y = col_red),
               alpha = .85, 
               fill = "blue", 
               color = "blue") +
  geom_col(aes(y = col_blue),
               alpha = .85, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  ylab("Monthly Change") +
  labs(title = "Wages and Hours",
       subtitle = "last 12 months") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
    strip.background = element_blank(),
    strip.placement = "inside",
    strip.text = element_text(size=15),
    panel.spacing = unit(0.2, "lines") ,
    panel.background=element_rect(fill="white")) +
  facet_wrap(~stat, shrink = FALSE)
```

### Wrangle the macro economic data

```{r}
fred_empl_data_wrangled <- 
  fred_empl_data  %>%
  select(date, 
         `Nat Rate Unemployment`, 
         `Unemployment Rate`,
         `Nonfarm Employment`,
         `Labor Force Part Rate`,
         `Employment-to-Pop Ratio`,
         `Nonfarm Employment`) %>% 
  mutate(`Nat Rate Unemployment` = na.locf(`Nat Rate Unemployment`, na.rm = F)) %>%
  mutate(`Unemployment Gap` = `Unemployment Rate` - `Nat Rate Unemployment`,
         `Employment Monthly Change` = `Nonfarm Employment` - lag(`Nonfarm Employment`))

```

### Create a recessions ggplot object

```{r}
recessions <- 
tribble(
  ~Peak, ~Trough,
  "1948-11-01", "1949-10-01",
  "1953-07-01", "1954-05-01",
  "1957-08-01", "1958-04-01",
  "1960-04-01", "1961-02-01",
  "1969-12-01", "1970-11-01",
  "1973-11-01", "1975-03-01",
  "1980-01-01", "1980-07-01",
  "1981-07-01", "1982-11-01",
  "1990-07-01", "1991-03-01",
  "2001-03-01", "2001-11-01",
  "2007-12-01", "2009-06-01"
  ) %>% 
  mutate(Peak = ymd(Peak),
         Trough = ymd(Trough))


recession_shade <- 
  geom_rect(data = recessions, 
            inherit.aes = F, 
            aes(xmin = Peak, 
                xmax = Trough, 
                ymin = -Inf, 
                ymax = +Inf), 
            fill = 'darkgray', 
            alpha = 0.5)
```


### Vis Showing Above or Below Trend

```{r}
fred_empl_data_wrangled %>% 
  mutate(ribbon_pink = 
           if_else(`Unemployment Rate` > `Nat Rate Unemployment`, 
                  `Unemployment Rate`, `Nat Rate Unemployment`),
         ribbon_blue = 
           if_else(`Unemployment Rate` < `Nat Rate Unemployment`, 
                  `Unemployment Rate`, `Nat Rate Unemployment`)) %>% 
  ggplot(aes(x = date, y = `Unemployment Rate`)) +
  geom_line(color = "black") +
  geom_line(linetype = 2, aes(y = `Nat Rate Unemployment`))  +
  recession_shade +
  geom_ribbon(aes(ymin = `Nat Rate Unemployment`, 
                  ymax = ribbon_pink), 
              fill = "blue", 
              alpha = 0.5)+
  geom_ribbon(aes(ymin = `Nat Rate Unemployment`,
                  ymax = ribbon_blue), 
              fill = "cornflowerblue", 
              alpha = 0.5)  +
  xlim(ymd("1980-01-01"), ymd("2018-03-01")) +
  labs(title = "Unemployment v. Nat Rate Unemploymnet",
       caption = "data source: fred",
       y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.caption  = element_text(hjust = 0)) 
```

### Grab the ADP and BLS Data from the Fred object

```{r}
adp_fred <-
  fred_empl_data  %>%
  select(date, ADP = `ADP Estimate`, BLS = `Nonfarm Employment`)  %>% 
    gather(report, month, -date) %>% 
    group_by(report) %>% 
    mutate(monthly_change = month - lag(month, 1))
```

```{r}
adp_fred %>% 
  filter(date >= "2017-01-01") %>%
  ggplot(aes(x = date, y = monthly_change, fill = report)) + 
  geom_col(position = "dodge") + 
  scale_fill_manual(values = c("blue","cornflowerblue")) +
  labs(title = "ADP Estimate v. BLS Report",
       caption = "data source: fred",
       y = "monthly change (thousands)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.caption  = element_text(hjust = 0)) +
  scale_x_date(breaks = scales::pretty_breaks(n = 16))
```

### Schedule email for BLS release dates

```{r}
# we will schedule this to run every friday in the RSC scheduler.
# On one Friday per month, when BLS releases jobs data, 
# this report will be emailed. 

release_dates <- 
tribble(
~"bls release date",  
"Jan. 05, 2018",
"Feb. 02, 2018",
"Mar. 09, 2018",
"Apr. 06, 2018",
"May. 04, 2018",
"Jun. 02, 2018",
"Jul. 06, 2018",
"Aug. 03, 2018",
"Sep. 07, 2018",
"Oct. 05, 2018",
"Nov. 02, 2018",
"Dec. 07, 2018") %>% 
mutate(`bls release date` = 
         ymd(parse_date_time(`bls release date`, "%m%d%Y"))) %>%
pull(`bls release date`)

 
if (today() %in% release_dates) {
  # send email if today is a release today
  rmarkdown::output_metadata$set(rsc_email_suppress_scheduled = FALSE)
} else {
  # don't send an email if today is not a release date
  rmarkdown::output_metadata$set(rsc_email_suppress_scheduled = TRUE)
}

```







