---
title: "Jobs Report"
output: html_notebook
---



```{r setup, include = FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
library(tidyverse)
library(tidyquant)
library(highcharter)
library(timetk)
library(tibbletime)
library(data.table)
library(geofacet)
library(scales)
library(readxl)
library(RcppRoll)
```


```{r}
states_hpi <- 
  read_excel("states-hpi.xls", 
        col_types = c("text", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "skip", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric",
        "numeric", "skip"), skip = 5) %>% 
  # rename usa hpi first
  rename(usa = `United States not seasonally adjusted`) %>% 
  slice(-517:-532) %>% 
  # parse the date with lubridate
  mutate(date = ymd(parse_date_time(as.character(Month), "%Y%m"))) %>% 
  select(date, everything(), -Month) %>% 
  gather(state, hpi, -date) %>% 
  group_by(state) %>%
  # use lag
  mutate(hpa12 = hpi/lag(hpi, 12) - 1,
         hpa1 = hpi/lag(hpi, 1) - 1,
         hpilag12 = lag(hpi, 12, fill = NA),
         # use RcppRoll functions
         hpimax12 = roll_max(hpi, 13, align = "right", fill = NA),
         hpimin12 = roll_min(hpi, 13, align = "right", fill = NA),
         year = year(date),
         month = month(date)) %>% 
  filter(month == 12 & year > 1989) %>% 
  select(year, state, hpa12)


my_states_grid <- states_hpi %>% 
  select(-Month) %>% 
  colnames()

my_state_codes <- 
  c(
  # nonfarm employment is NA for seasonally adjusted, change to NAN for not adjusted. weird
  paste0(my_states_grid,"NA"),      
  # resident population (annual)
  paste0(my_states_grid,"POP"),
  # private building permits (monthly, NSA)
  paste(my_states_grid, "BPPRIV", sep = "")  
           )

state_emp_pop_perm_data <- tq_get(my_state_codes,
                                  get = "economic.data", 
                                  from = "2017-06-01")

recessions.df = read.table(textConnection(
  "Peak, Trough
  1948-11-01, 1949-10-01
  1953-07-01, 1954-05-01
  1957-08-01, 1958-04-01
  1960-04-01, 1961-02-01
  1969-12-01, 1970-11-01
  1973-11-01, 1975-03-01
  1980-01-01, 1980-07-01
  1981-07-01, 1982-11-01
  1990-07-01, 1991-03-01
  2001-03-01, 2001-11-01
  2007-12-01, 2009-06-01"), sep=',',
  colClasses = c('Date', 'Date'), header = TRUE)

```

Import non farm payroll data

```{r}
fred_codes <- c(
            "CIVPART", # civilian labor force pariticipation rate
            "EMRATIO", # employment-to-population ratio
            "NROU",  # estimate of natural rate of unemployment from U.S. CBO
            "PAYEMS", # nonfarm payroll employment
            "UNRATE") # unemployment rate
            
column_names <- c(
            "date",
            "Labor Force Part Rate",
            "Employment-to-Pop Ratio",
            "Nat Rate Unemployment",
            "Nonfarm Employment",
            "Unemployment Rate")


# download data via FRED 
fred_empl_data <- tq_get(fred_codes,                         # get selected symbols
            get = "economic.data",             # use FRED
            from = "1980-01-01")  

fred_empl_data_wrangled <- 
  fred_empl_data %>% 
  spread(symbol, price) %>% 
  `colnames<-`(column_names) %>% 
 # gather(set, value, -date)
  # make nat rate unemployment monthly by filling in the NAs
  mutate(`Nat Rate Unemployment` = na.locf(`Nat Rate Unemployment`, na.rm = F)) %>%
  
  mutate(`Unemployment Gap` = `Unemployment Rate` - `Nat Rate Unemployment`,
         `Employment Monthly Change` = `Nonfarm Employment` - lag(`Nonfarm Employment`),
         # create indicators for shaded plot
         up = if_else(`Unemployment Rate` > `Nat Rate Unemployment`, 
                      `Unemployment Rate`, 
                      `Nat Rate Unemployment`),
         down = if_else(`Unemployment Rate` < `Nat Rate Unemployment`, 
                        `Unemployment Rate`, 
                        `Nat Rate Unemployment`))
```


```{r}
fred_empl_data_wrangled %>%
  mutate(col_blue = 
           if_else(`Employment Monthly Change` > 0, 
                  `Employment Monthly Change`, NULL),
         col_red = 
           if_else(`Employment Monthly Change` < 0, 
                  `Employment Monthly Change`, NULL)) %>%
  ggplot(aes(x = date)) +
  geom_col(aes(y = col_red),
               alpha = .85, 
               fill = "pink", 
               color = "pink") +
  geom_col(aes(y = col_blue),
               alpha = .85, 
               fill = "cornflowerblue", 
               color = "cornflowerblue") +
  ylab("Monthly Change") +
  ggtitle("Monthly Private Employment Change") +
  scale_x_date(date_breaks = "2 years", 
               date_labels = "%Y") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10),
                     labels = scales::comma) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
  
```


```{r}
fred_empl_data_wrangled %>% 
  mutate(ribbon_pink = 
           if_else(`Unemployment Rate` > `Nat Rate Unemployment`, 
                  `Unemployment Rate`, `Nat Rate Unemployment`),
         ribbon_blue = 
           if_else(`Unemployment Rate` < `Nat Rate Unemployment`, 
                  `Unemployment Rate`, `Nat Rate Unemployment`)) %>%
  ggplot(aes(x = date, y = `Unemployment Rate`)) +
  geom_rect(data = recessions.df, 
            inherit.aes = F, 
            aes(xmin = Peak, xmax = Trough, ymin = -Inf, ymax = +Inf), 
            fill = 'darkgray', 
            alpha = 0.5)  +
  geom_line(color = "black") +
  geom_line(linetype = 2, aes(y = `Nat Rate Unemployment`)) +
  geom_ribbon(aes(ymin = `Nat Rate Unemployment`, 
                  ymax = ribbon_pink), 
              fill = "pink", 
              alpha = 0.5)+
  geom_ribbon(aes(ymin = `Nat Rate Unemployment`,
                  ymax = ribbon_blue), 
              fill = "cornflowerblue", 
              alpha = 0.5)  +
  xlim(ymd("1980-01-01"), ymd("2018-03-01"))
```


Visualize employment data

```{r}

```


Import and wrangle the labor force participation data

```{r}
labor_for_part_data <- 
  fread("https://download.bls.gov/pub/time.series/ln/ln.data.1.AllData") %>%
  filter(series_id == "LNS11300060" |
        series_id == "LNS11300061" | 
        series_id == "LNS11300062") %>% 
  mutate(date = ymd(parse_date_time(paste0(year, period), "%Y%m")),
         dem = case_when(series_id == "LNS11300060" ~ "All",
                         series_id == "LNS11300061" ~ "Men",
                         series_id == "LNS11300062" ~ "Women"),
         value = as.numeric(value)) %>% 
  select(date, everything(), -footnote_codes, -year, -period, -series_id)
```

Graph over time with ggplot.

```{r}

labor_for_part_data %>% 
  group_by(dem) %>%
  #filter(date >= "1980-01-01") %>% 
  ggplot(aes(x = date, y = value, color = dem, label = dem)) +
  geom_rect(data = recessions.df, 
            inherit.aes = F, 
            aes(xmin = Peak, 
                xmax = Trough, 
                ymin = -Inf, 
                ymax = +Inf), 
            fill = 'darkgray', alpha = 0.5) +
  geom_line()  +
  scale_x_date(date_breaks = "5 years", 
               date_labels = "%Y") +
  #limits = c(ymd("1980-01-01"), ymd("2018-02-01"))) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10), 
                     labels = function(x) paste0(x, "%"), 
                     sec.axis = dup_axis()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption=element_text(hjust=0)) +
  ylab("") +
  xlab("") +
  ggtitle("Labor Force Participation Rate: Prime Working Age (25-54)", 
          subtitle = "seasonally adjusted") +
  labs(caption = "based on work by the inestimable @lenkiefer") 
  
``

