---
title: "Jobs Report End to End"
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include = FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(tidyquant)
#devtools::install_github("jimhester/vroom")
# library(vroom)
```

### Import data via FRED and BLS

One of the great things about R is that packages have been built to help us with a lot of tasks. 

We will use `tq.get()` from the `tidyquant` package to retrieve data from FRED. First, we need to indicate what data we want.

```{r}
fred_codes <- c(
            "CIVPART", 
            "EMRATIO",
            "NROU",  
            "UNRATE", 
            "CES0500000003",
            "AWHAETP",
            "NPPTTL",
            "PAYEMS", 
            "USCONS",
            "USTRADE",
            "USPBS",
            "MANEMP",
            "USFIRE",
            "USMINE",
            "USEHS",
            "USWTRADE",
            "CES4349300001",
            "USINFO",
            "USLAH",
            "USGOVT",
            "USSERV")
          
column_names <- c(
            "date",
            "Labor Force Part Rate",
            "Employment-to-Pop Ratio",
            "Nat Rate Unemployment",
            "Unemployment Rate",
            "Avg Hourly Earnings",
            "Avg Weekly Hours",
            "ADP Estimate",
            "Nonfarm Employment",
            "Construction",
            "Retail/Trade",
            "Prof/Bus Serv",
            "Manufact",
            "Financial",
            "Mining",
            "Health Care",
            "Wholesale Trade",
            "Transportation",
            "Info Sys",
            "Leisure",
            "Gov",
            "Other Services")

library(tidyquant)
# download data via FRED by using tidyquant
fred_empl_data <- 
  tq_get(fred_codes,                         
         get = "economic.data",             
         from = "1980-01-01")
fred_empl_data %>% 
  na.omit()
```

Let's go get some BLS data that for whatever reason FRED doesn't include. In this case, labor force participation rate for the working age population age 25-54, and then for men and women. 


```{r}
# download data direct from BLS
# use the 'import dataset' button in RStudio to experiment with how to get this data
# into a good, importable format
bls_labor_for_data <- 
  read_delim("https://download.bls.gov/pub/time.series/ln/ln.data.1.AllData", 
             "\t", 
             escape_double = FALSE, 
             col_types = cols(
               value = col_character(),
               footnote_codes = col_skip()),
             trim_ws = TRUE,
             progress = FALSE)


```

We just imported 6.3 million rows and every time series that BLS has available. But we want to work with just three. Let use `filter` from `dplyr` by series_id to get just our desired rows, then use `parse_date_time` from the `lubridate` package to create a nicer date. We will use `mutate` and `case_when` from `dplyr` to create a column of demographic categories.

```{r}


bls_labor_wrangled <- 
  bls_labor_for_data %>% 
  filter(series_id == "LNS11300060"|
        series_id == "LNS11300061" | 
        series_id == "LNS11300062") %>% 
  mutate(date = ymd(parse_date_time(paste0(year, period), "%Y%m")),
         dem = case_when(series_id == "LNS11300060" ~ "All",
                         series_id == "LNS11300061" ~ "Men",
                         series_id == "LNS11300062" ~ "Women"),
         value = as.numeric(value)) %>%
  filter(date >= "1980-01-01") %>% 
  select(date, everything(), -year, -period, -series_id)

```

### API

We might want to use an API like Energy Information Administration (www.eia.gov). All the data is public available through their OpenData platform. Anyone can acquire a free token and access all of their public data sets.

```{r}
library(rjson)
library(RJSONIO)

# code for gas storage all USA
code <-   "NG.NW2_EPG0_SWO_R48_BCF.W"

fromJSON(paste0(
    "http://api.eia.gov/series/?api_key=B0EDABE19140380774029141270387BE&series_id=",
    code,
    "&format=json")) %>%  
  pluck("series", 1, "data") %>%
  transpose() %>% 
  `names<-`(c("date", "value")) %>% 
  simplify_all() %>% 
  as_tibble() %>% 
  mutate(date = ymd(date)) %>% 
  arrange(date)
```



```{r}

eia_df_gasstorage <-
tribble(
  ~ticker, ~location,
  "NG.NW2_EPG0_SWO_R48_BCF.W", "US",
  "NG.NW2_EPG0_SWO_R31_BCF.W", "East",
  "NG.NW2_EPG0_SWO_R32_BCF.W", "Midwest",
  "NG.NW2_EPG0_SWO_R35_BCF.W", "Pacific",
  "NG.NW2_EPG0_SWO_R33_BCF.W", "South"
  ) 

# best is to wrap to function when done
eia_grab_func <- function(code) {
  fromJSON(paste0(
    "http://api.eia.gov/series/?api_key=B0EDABE19140380774029141270387BE&series_id=",
    code,
    "&format=json")) %>%  
  pluck("series", 1, "data") %>%
  transpose() %>% 
  `names<-`(c("date", "value")) %>% 
  simplify_all() %>% 
  as_tibble() %>% 
  mutate(date = ymd(date)) %>% 
  arrange(date)
}

eia_df_gasstorage %>%  
  mutate(data = map(ticker, eia_grab_func)) %>% 
  unnest() %>% 
  head()

```


### Visualizations

We will use just the BLS data from here.


Now we visualize our three data sets over time. `ggplot` can color each time series uniqely since we tell it to `group_by` the `dem` column. We add our recessions object that was created in the above code chunk.

```{r, warning= FALSE}

bls_line_chart <- bls_labor_wrangled %>% 
  group_by(dem) %>% 
  ggplot(aes(x = date, 
             y = value, 
             color = dem, 
             label = dem)) +
  geom_line()  +
  theme_minimal() +
  xlim(ymd("1980-01-01"), Sys.Date()) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10), 
                     labels = function(x) paste0(x, "%"), 
                     sec.axis = dup_axis()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption=element_text(hjust=0)) +
  ylab("") +
  xlab("") +
  ggtitle("Labor Force Participation Rate: Prime Working Age (25-54)", 
          subtitle = "seasonally adjusted") +
  labs(caption = "based on work by FMAC's @lenkiefer") 

bls_line_chart
```

```{r}
# a quick ggplot recessions data code flow
recessions_tribble <- tribble(
  ~Peak, ~Trough,
  "1948-11-01", "1949-10-01",
  "1953-07-01", "1954-05-01",
  "1957-08-01", "1958-04-01",
  "1960-04-01", "1961-02-01",
  "1969-12-01", "1970-11-01",
  "1973-11-01", "1975-03-01",
  "1980-01-01", "1980-07-01",
  "1981-07-01", "1982-11-01",
  "1990-07-01", "1991-03-01",
  "2001-03-01", "2001-11-01",
  "2007-12-01", "2009-06-01") %>% 
  mutate(Peak = ymd(Peak),
         Trough = ymd(Trough))

# can save ggplot variables
# geom_rect creates that object
recession_shade <- 
  geom_rect(data = recessions_tribble, 
            inherit.aes = F, 
            aes(xmin = Peak, 
                xmax = Trough, 
                ymin = -Inf, 
                ymax = +Inf), 
            fill = 'darkgray', 
            alpha = 0.4)

bls_line_chart + recession_shade
```


### Highcharter with tidy objects

```{r}
library(highcharter)
hchart(bls_labor_wrangled, 
       type = 'line', 
       hcaes(y = value,
             x = date,
             group = dem)) %>% 
  hc_title(text = "Labor Force Participation (age 25-54") %>% 
      hc_xAxis(type = 'datetime',
                 plotBands = list(
                   list(
                     label = list(text = "Bush Sr."),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('1990-07-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('1991-03-01', tz = 'UTC'))
                     ),
                   list(
                     label = list(text = "Dot Com"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('2001-03-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('2001-11-01', tz = 'UTC'))
                     ),
                   list(
                     label = list(text = "Fin Crisis"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('2007-12-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('2009-06-01', tz = 'UTC'))
                     ),
                   list(
                     label = list(text = "Volker Fed"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('1981-07-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('1982-11-01', tz = 'UTC'))
                     )
                   )
               ) %>% 
  hc_add_theme(hc_theme_flat()) %>%
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_exporting(enabled = TRUE)
```

### Convert to xts

A different data format, sometimes needed for visualizations or models. We will convert our data to an `xts` object. XTS stands for extensible time series. Note there is not date 'column', it's a date index. If we remove that index, it's no longer an `xts` object, it reverts to a matrix.

Our conversion is done with the `tk_xts` function from the `timetk` package.

```{r}
library(timetk)

bls_labor_xts <- 
  bls_labor_wrangled%>% 
  group_by(dem) %>% 
  spread(dem, value) %>% 
  tk_xts(date_var = date)
```

### Visualize xts with highcharter

highcharter is an R hook into the highcarts javascript library. Similar js hooks are available through the `dygraphs` and `plotly` packages.

```{r}
library(highcharter)


highchart(type = 'stock')%>%
  hc_add_series(bls_labor_xts$All, name = "All") %>%
  hc_add_series(bls_labor_xts$Men, name = "Men") %>%
  hc_add_series(bls_labor_xts$Women, name = "Women") %>%
  hc_title(text = "Labor Force Participation (age 25-54") %>% 
      hc_xAxis(type = 'datetime',
                 plotBands = list(
                   list(
                     label = list(text = "Bush Senior"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('1990-07-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('1991-03-01', tz = 'UTC'))
                     ),
                   list(
                     label = list(text = "Internet Bubble"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('2001-03-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('2001-11-01', tz = 'UTC'))
                     ),
                   list(
                     label = list(text = "Financial Crisis"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('2007-12-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('2009-06-01', tz = 'UTC'))
                     ),
                   list(
                     label = list(text = "Volker Fed"),
                     color = "rgba(100, 0, 0, 0.1)",
                     from = datetime_to_timestamp(as.Date('1981-07-01', tz = 'UTC')),
                     to = datetime_to_timestamp(as.Date('1982-11-01', tz = 'UTC'))
                     )
                   )
               ) %>% 
  hc_navigator(enabled = FALSE) %>% 
  hc_scrollbar(enabled = FALSE) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_exporting(enabled = TRUE)
```


### Convert to TS

Now let's do some modeling and forecasting using the `forecast` package. The `forecast` functions want data in yet another format, that of `ts`. So, we need to convert again with the `tk_ts` function from `timetk`. We will go to quarterly data.

```{r}
library(timetk)

bls_all_ts <- 
  bls_labor_wrangled %>%
  filter(date >= "1990-01-01") %>%
  filter(dem == "All") %>% 
  tq_transmute(mutate_fun = to.period, period = "quarters") %>% 
    tk_ts(select = -date, start = c(1990, 3), freq = 4)
```



### Model and forecast as arima, then ets, then bats. 

We will use the build in models from the forecast package but can use custom or other models too.

### Auto Arima model

```{r}

library(forecast)
auto.arima(bls_all_ts) %>% 
  forecast() %>% 
  hchart()
```

### ETS model
```{r}
ets(bls_all_ts, damped = TRUE) %>% 
  forecast()%>% 
  hchart()
```

### Bats model

```{r}
bats(bls_all_ts) %>% 
  forecast()%>% 
  hchart()
```

### N Models, one piped workflow

First create a list of models, and nest our `ts` data as list inside those lists. 

```{r}

fcast_model_list <- list(
   naive = list(
      y = bls_all_ts
    ),  
   rwf = list(
      y = bls_all_ts,
      drift = TRUE
    ),
    auto.arima = list(
        y = bls_all_ts
        ),
    ets = list(
        y = bls_all_ts,
        damped = TRUE
    )
)
```

Use `enframe` from the `tibble` package to turn our list of functions into a data frame.

```{r}


fcast_model_tbl <- 
  enframe(fcast_model_list, 
          name = "func", 
          value = "params")

fcast_model_tbl
```



### Sweep the data

The `sweep` package was built to tidy up time series models and forecasts. Let's use the `sw_sweep` function.

```{r}
library(sweep)

fcast_model_tbl %>%
mutate(modeled = invoke_map(func, params)) %>% 
mutate(forecasted = map(modeled, forecast, h = 10)) %>% 
mutate(swept = map(forecasted, 
                   sw_sweep, 
                   fitted = FALSE, 
                   timetk_idx = TRUE, 
                   rename_index = "date")) %>% 
unnest(swept) 

```

### It's tidy, let's use ggplot for visualizations


```{r}

fcast_model_tbl %>%
mutate(modeled = invoke_map(func, params)) %>% 
mutate(forecasted = map(modeled, forecast, h = 10)) %>% 
mutate(swept = map(forecasted, 
                   sw_sweep, 
                   fitted = FALSE, 
                   timetk_idx = TRUE, 
                   rename_index = "date")) %>% 
unnest(swept) %>% 
ggplot(aes(x = date, 
           y = value, 
           color = key, 
           group = func)) +
    geom_ribbon(aes(ymin = lo.95, 
                    ymax = hi.95), 
                fill = "pink", 
                color = NA, 
                size = 0) +
    geom_ribbon(aes(ymin = lo.80, 
                    ymax = hi.80), 
                fill = "cornflowerblue", 
                color = NA, 
                size = 0, 
                alpha = 0.5) +
    geom_line() +
    facet_wrap(~func, nrow = 3) +
  labs(title = "Labor Force Part Forecasts",
       x = "",
       y = "") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_x_date(breaks = scales::pretty_breaks(n = 15)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) 



```


```{r}

fcast_model_tbl %>%
mutate(modeled = invoke_map(func, params)) %>% 
mutate(forecasted = map(modeled, forecast, h = 10)) %>% 
mutate(swept = map(forecasted, 
                   sw_sweep, 
                   fitted = FALSE, 
                   timetk_idx = TRUE, 
                   rename_index = "date")) %>% 
unnest(swept) %>% 
ggplot(aes(x = date, 
           y = value, 
           color = key, 
           group = func)) +
    # geom_ribbon(aes(ymin = lo.95, 
    #                 ymax = hi.95), 
    #             fill = "pink", 
    #             color = NA, 
    #             size = 0) +
    # geom_ribbon(aes(ymin = lo.80, 
    #                 ymax = hi.80), 
    #             fill = "cornflowerblue", 
    #             color = NA, 
    #             size = 0, 
    #             alpha = 0.5) +
    geom_line(aes(color = func)) +
    #facet_wrap(~func, nrow = 3) +
  labs(title = "Labor Force Part Forecasts",
       x = "",
       y = "") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_x_date(breaks = scales::pretty_breaks(n = 15)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) 



```

Save our swept object.

```{r}

fcast_swept <- fcast_model_tbl %>%
mutate(modeled = invoke_map(func, params)) %>% 
mutate(forecasted = map(modeled, forecast, h = 10)) %>% 
mutate(swept = map(forecasted, 
                   sw_sweep, 
                   fitted = FALSE, 
                   timetk_idx = TRUE, 
                   rename_index = "date")) %>% 
unnest(swept)

fcast_swept %>% 
  head()
```

To color different forecasts by different colors, first create two new columns. One to hold the actual values, called `actual_col` and one to hold just the forecast values, called `forecast_col`. 

```{r}
fcast_swept %>% 
 mutate(forcast_col = case_when(key == "forecast" ~ value),
        actual_col = case_when(key == "actual" ~ value)) %>% 
  dplyr::select(date, func, key, actual_col,  forcast_col)
```

Now tell `ggplot` to chart two lines from two columns. One line is from the actual values. The second line is from the forecasts and so will have N values and N colors.

```{r}
fcast_swept %>% 
 mutate(forcast_col = case_when(key == "forecast" ~ value),
        actual_col = case_when(key == "actual" ~ value)) %>% 
  dplyr::select(date, func, key, actual_col,  forcast_col) %>% 
  ggplot(aes(x = date)) +
  geom_line(aes(y = actual_col), color = "cornflowerblue") +
  geom_line(aes(y = forcast_col, color = func, label = func)) +
  labs(title = "Labor Force Part Forecasts",
       x = "",
       y = "") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  scale_x_date(breaks = scales::pretty_breaks(n = 15)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5)) 

```

```{r}

fcast_swept_hc <- 
  fcast_swept %>% 
 dplyr::select(func, date, value) %>% 
  spread(func, value) %>% 
  tk_xts(date_var =  date)


highchart(type = 'stock')%>%
  hc_add_series(fcast_swept_hc$auto.arima, name = "arima") %>%
  hc_add_series(fcast_swept_hc$ets, name = "ets") %>%
  hc_add_series(fcast_swept_hc$naive, name = "naive") %>%
  hc_add_series(fcast_swept_hc$rwf, name = "drift")

```

<!-- ### Rolling Standard Deviation by group -->

<!-- There was a question about rolling volatility by group.  -->

<!-- ```{r} -->

<!-- # The symbols vector holds our tickers.  -->
<!-- symbols <- c("SPY","EFA", "IJS", "EEM","AGG") -->

<!--  etf_returns <- -->
<!--    symbols %>%  -->
<!--    # get prices from yahoo -->
<!--    tq_get(get  = "stock.prices", -->
<!--            from = "2011-12-31", -->
<!--            to   = "2017-12-31") %>%  -->
<!--     group_by(symbol) %>% -->
<!--     # convert to monthly returns -->
<!--     tq_transmute(select     = adjusted,  -->
<!--                  mutate_fun = periodReturn,  -->
<!--                  period     = "monthly",  -->
<!--                  col_rename = "monthly_returns") -->
<!--  etf_returns -->


<!-- ``` -->

<!-- Have a look at the data. It's monthly returns for each ETF, grouped by the ETF symbol. Now we can run calculations on a group by group basis.  -->

<!-- Two ways to do so:  -->

<!-- 1) use the built in functionality from `tidyquant`, which wraps `rollapply` and the `sd` function and allows us to apply to a tidy tibble. -->

<!-- ```{r} -->
<!-- # width = the rolling window -->
<!-- etf_returns %>%  -->
<!--   tq_mutate(mutate_fun = rollapply, -->
<!--             width = 6, -->
<!--             FUN = sd, -->
<!--             col_rename = "rolling_sd") %>% -->
<!--   select(date, rolling_sd) -->

<!-- ``` -->

<!-- That method works very well. Let's see how to do the same with `tibbletime` -->

<!-- First create our own rolling function using `rollify` from the `tibbletime` package. -->

<!-- ```{r}   -->
<!-- library(tibbletime) -->
<!-- sd_roll_6 <-  -->
<!--   rollify(sd, window = 6) -->
<!-- ``` -->

<!-- Now convert `etf_returns` to a `tibbletime` object and apply that function. -->

<!-- ```{r} -->

<!-- etf_returns %>% -->
<!--   as_tbl_time(index = date) %>%  -->
<!--   mutate(sd = sd_roll_6(monthly_returns)) %>%  -->
<!--   select(-monthly_returns) -->


<!-- ``` -->

<!-- Those two methods got us the same result, which one is better? Well, the first method is a bit easier because we didn't need to create our own rolling function. The second method is more flexible - we can create whatever rolling function we want and apply it to a tibbletime object.  -->



