---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidyquant)
library(rvest)
```


```{r}
# Read html
html <- read_html("vg_info.html")
# Extract fund names
vanguard_fund_names <- html %>%
  html_nodes(".productName a") %>%
  html_text()
  
write_rds(vanguard_fund_names, "vanguard_fund_names.RDS")

# Extract tickers
vanguard_tickers <- html %>%
  html_nodes(".fixedCol+ .ng-binding") %>%
  html_text()

write_rds(vanguard_tickers, "vanguard_tickers.RDS")

# Extract asset classes
vanguard_asset_class <- html %>%
  html_nodes(".assetClass") %>%
  html_text() %>%
  sapply(., function(x) x[!str_detect(x, "\t")]) %>%
  unlist()

write_rds(vanguard_asset_class, "vanguard_asset_class.RDS")

vanguard_ticker_fund_asset_class_tibble <- 
  tibble(
    ticker = vanguard_tickers,
    fund = vanguard_fund_names,
    asset_class = vanguard_asset_class
      
  )

write_rds(vanguard_ticker_fund_asset_class_tibble, "vanguard_ticker_fund_asset_class_tibble.RDS")

library(pins)


pins::board_register(server = "https://colorado.rstudio.com/rsc", key = "fAEzqlEWVZM8GEQAkoQiKTOYgMXrAvZ2", board = "rsconnect")

pin(vanguard_ticker_fund_asset_class_tibble, description = "Vanguard tickers, funds and asset classes", board = "rsconnect")


# Extract expense ratios
expense_ratio <- html %>%
  html_nodes(".expenseRatio") %>%
  html_text()  %>%
  sub("%", "", .) %>%
  as.numeric() %>%
  sapply(., '/', 100) %>%
  sapply(., function(x) x[!is.na(x)]) %>%
  unlist()
# Extract price
price <- html %>%
  html_nodes(".expenseRatio+ .ng-binding") %>%
  html_text()  %>%
  sub("[$]", "", .) %>%
  as.numeric()
# Extract sec yield
sec_yield <- html %>%
  html_nodes(".secYield") %>%
  html_text()
sec_yield_clean <- sec_yield[!str_detect(sec_yield, "SEC")] %>%
  str_replace_all(., "\n", "") %>%
  str_replace_all(., "\t", "") %>%
  str_replace_all(., "창\u0080\u0094", NA_character_)
# Extract ytd returns
ytd <- html %>%
  html_nodes(".secYield+ .ng-binding") %>%
  html_text() %>%
  sub("%", "", .) %>%
  as.numeric() %>%
  sapply(., '/', 100)
# Extract one yr returns
one_yr <- html %>%
  html_nodes(".ng-binding:nth-child(11)") %>%
  html_text() %>%
  str_replace_all(., "창\u0080\u0094", NA_character_) %>%
  sub("%", "", .) %>%
  as.numeric() %>%
  sapply(., '/', 100)
# Extract five yr returns
five_yr <- html %>%
  html_nodes(".ng-binding:nth-child(12)") %>%
  html_text() %>%
  str_replace_all(., "창\u0080\u0094", NA_character_) %>%
  sub("%", "", .) %>%
  as.numeric() %>%
  sapply(., '/', 100)
# Extract ten yr yields
ten_yr <- html %>%
  html_nodes(".ng-binding:nth-child(13)") %>%
  html_text() %>%
  str_replace_all(., "창\u0080\u0094", NA_character_) %>%
  sub("%", "", .) %>%
  as.numeric() %>%
  sapply(., '/', 100)

# Extract yield since inception
since <- html %>%
  html_nodes(".right:nth-child(14)") %>%
  html_text() %>%
  str_replace_all(., "\n", "") %>%
  str_replace_all(., "\t", "") %>%
  str_split(., "[(]") %>%
  lapply(., head, 1) %>%
  unlist() %>%
  sub("%", "", .) %>%
  as.numeric() %>%
  sapply(., '/', 100)

# Extract date of inception
inception <-  html %>%
  html_nodes(".right:nth-child(14)") %>%
  html_text() %>%
  str_replace_all(., "\n", "") %>%
  str_replace_all(., "\t", "") %>%
  str_split(., "[(]") %>%
  lapply(., tail, 1) %>%
  str_replace_all(., "[)]", "") %>%
  unlist() %>%
  as.Date(., "%m/%d/%Y")
# Combine into one data frame
fund_data <- data.frame(fund_names, ticker, asset_class,
                        expense_ratio, price, sec_yield_clean, 
                        ytd, one_yr, five_yr, ten_yr, since, inception,
                        stringsAsFactors = FALSE)
# Drop duplicate rows
fund_data <- fund_data %>%
  distinct()

write.csv(fund_data,
          "fund_data.csv",
          row.names = FALSE)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

