---
title: "R Notebook"
output: html_notebook
---
 
```{r setup, include=FALSE, warning=FALSE, message=FALSE}

library(reticulate)
library(tidyverse)
library(tidyquant)
library(janitor)
library(collapsibleTree)
library(formattable)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r}

iefn_holdings <- read_csv("https://www.ishares.com/us/products/292421/fund/1467271812596.ajax?fileType=csv&fileName=IEFN_holdings&dataType=fund", 
    col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "iefn", 
         download_date = ymd(Sys.Date())) 

iehs_holdings <- read_csv("https://www.ishares.com/us/products/292422/fund/1467271812596.ajax?fileType=csv&fileName=IEHS_holdings&dataType=fund", 
    col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "iehs", 
         download_date = ymd(Sys.Date()))

ieih_holdings <- read_csv("https://www.ishares.com/us/products/292423/fund/1467271812596.ajax?fileType=csv&fileName=IEIH_holdings&dataType=fund", 
    col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "ieih", 
         download_date = ymd(Sys.Date()))

ieme_holdings <- read_csv("https://www.ishares.com/us/products/292424/fund/1467271812596.ajax?fileType=csv&fileName=IEME_holdings&dataType=fund", 
    col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "ieme", 
         download_date = ymd(Sys.Date()))

ietc_holdings <- read_csv("https://www.ishares.com/us/products/292425/fund/1467271812596.ajax?fileType=csv&fileName=IETC_holdings&dataType=fund", 
    col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "ietc", 
         download_date = ymd(Sys.Date()))

iecs_holdings <- read_csv("https://www.ishares.com/us/products/292414/fund/1467271812596.ajax?fileType=csv&fileName=IECS_holdings&dataType=fund", 
    col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "iecs", 
         download_date = ymd(Sys.Date()))


iede_holdings <- read_csv("https://www.ishares.com/us/products/292415/fund/1467271812596.ajax?fileType=csv&fileName=IEDE_holdings&dataType=fund", 
    col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "iede", 
         download_date = ymd(Sys.Date()))

msci_usa_holdings <- read_csv("https://www.ishares.com/us/products/239693/ishares-msci-usa-etf/1467271812596.ajax?fileType=csv&fileName=EUSA_holdings&dataType=fund", col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "msci_usa", 
         download_date = ymd(Sys.Date()))

msci_em_holdings <- read_csv("https://www.ishares.com/us/products/239637/ishares-msci-emerging-markets-etf/1467271812596.ajax?fileType=csv&fileName=EEM_holdings&dataType=fund", col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "msci_em", 
         download_date = ymd(Sys.Date()))
```

Vanguard manual - VG makes it tough, have to go click the CSV then import. I'm working on a script for this.

```{r}
vti_total_mkt_holdings <- read_csv("Top_holdings_Total_Stock_Market_ETF.csv", 
    col_types = cols(COUNTRY = col_skip(), 
        `SECURITY DEPOSITORY RECEIPT TYPE` = col_skip(), 
        SEDOL = col_skip(), X1 = col_skip()), 
    skip = 7)
```

Clean up the column names and determine how many tickers we have

```{r}
vti_total_mkt_holdings %>% 
  janitor::clean_names() %>% 
  mutate(ticker = str_replace(ticker, "BRK.B", "BRK-B")) %>% 
  distinct(ticker)
```

About 3,500 tickers. For today's purposes, let's take the top 1000 holdings. Even that will be a hefty data import but if you work in industry, this shouldn't be an issue for you as you probably have a feed. 

```{r}
library(readr)

vgt_holdings <- read_csv("Top_holdings_Information_Technology_ETF.csv", 
    col_types = cols(`SECURITY DEPOSITORY RECEIPT TYPE` = col_skip(), 
        SEDOL = col_skip(), X1 = col_skip()), 
    skip = 7) %>% 
  mutate(fund = "vgt",
         download_date = ymd(Sys.Date())) %>% 
  rename(Name = HOLDINGS)


```


```{r}
vgk_holdings <- 
  read_csv("Top_holdings_FTSE_Europe_ETF.csv", 
           col_types = cols(`SECURITY DEPOSITORY RECEIPT TYPE` = col_skip(), SEDOL = col_skip(), X1 = col_skip()), 
           skip = 6)%>% 
  mutate(fund = "vgk",
         download_date = ymd(Sys.Date())) %>% 
  rename(Name = HOLDINGS)
```


```{r}
ieme_tickers <- 
ieme_holdings %>% 
  pull(Ticker)

ieme_prices <- 
ieme_tickers %>% 
  tq_get(get = "stock.prices",
         from = "2010-01-01")

colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')

ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)
library(plotly)
ieme_prices %>% 
  group_by(symbol) %>% 
  mutate(year = year(date),
         daily_return = close/lag(close) -1) %>% 
  group_by(symbol, year) %>% 
         summarise(yearly_sd = sd(daily_return, na.rm = T), 
                   yearly_mean = mean(daily_return, na.rm = T)) %>% 
  left_join(ieme_holdings %>% select(Ticker, Name), by = c("symbol" = "Ticker")) %>% 
  plot_ly(x = ~yearly_sd, y = ~yearly_mean, color = ~Name, size = ~yearly_mean/yearly_sd, colors = colors, frame = ~year,
        type = 'scatter', mode = 'markers', sizes = c(5,50),
        marker = list(symbol = 'circle', sizemode = 'diameter',
                      line = list(width = 2, color = '#FFFFFF')),
        text = ~paste('ticker:', symbol, '<br>name:', Name, '<br>return:', scales::percent(yearly_mean),
                      '<br>annualized sd:', scales::percent(yearly_sd))) %>% 
    layout(xaxis = list(tickformat = "%", title = " sd"),
           yaxis = list(tickformat = "%", title = " return"),
           plot_bgcolor = "#000000",
         paper_bgcolor = "#000000",
         font = list(color="#ffffff"),
      showlegend = F) %>% 
  animation_opts(1000) %>% 
  animation_button(visible = T) %>% 
  config(displayModeBar = F)
```

```{r}
library(plotly)
holdings_visualizer_function <- function(fund_tibble){

tickers <- 
fund_tibble %>% 
  pull(Ticker)

prices <- 
tickers %>% 
  tq_get(get = "stock.prices",
         from = "2010-01-01")

colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')

ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)


prices %>% 
  group_by(symbol) %>% 
  mutate(year = year(date),
         daily_return = close/lag(close) -1) %>% 
  group_by(symbol, year) %>% 
         summarise(yearly_sd = sd(daily_return, na.rm = T), 
                   yearly_mean = mean(daily_return, na.rm = T)) %>% 
  left_join(fund_tibble %>% select(Ticker, Name), by = c("symbol" = "Ticker")) %>% 
  plot_ly(x = ~yearly_sd, y = ~yearly_mean, color = ~Name, size = ~yearly_mean/yearly_sd, colors = colors, frame = ~year,
        type = 'scatter', mode = 'markers', sizes = c(5,50),
        marker = list(symbol = 'circle', sizemode = 'diameter',
                      line = list(width = 2, color = '#FFFFFF')),
        text = ~paste('ticker:', symbol, '<br>name:', Name, '<br>return:', scales::percent(yearly_mean),
                      '<br>annualized sd:', scales::percent(yearly_sd))) %>% 
    layout(xaxis = list(tickformat = "%", title = " sd"),
           yaxis = list(tickformat = "%", title = " return"),
           plot_bgcolor = "#000000",
         paper_bgcolor = "#000000",
         font = list(color="#ffffff"),
      showlegend = F) %>% 
  animation_opts(1000) %>% 
  animation_button(visible = T) %>% 
  config(displayModeBar = F)
}


holdings_visualizer_function(ieme_holdings)
```


```{r}
p_dat <- 
  stock_p %>% 
  left_join(stock_index) %>% 
  select(symbol,company,sector,date,weight,adjusted,ret,cumulative_ret) %>% 
  mutate(year = year(date)) %>% 
  group_by(symbol,sector,year) %>% 
  mutate(cumulative_ret = with_order(date,cumprod,1+ret)-1, 
         avg_sd = sd(ret,na.rm=T)*sqrt(252)) %>% 
  filter(date == max(date,na.rm=T)) %>% 
  ungroup()

colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')

ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)


p_dat %>% 
  plot_ly(x = ~avg_sd, y = ~cumulative_ret, color = ~sector, size = ~cumulative_ret/avg_sd, colors = colors, frame = ~year,
        type = 'scatter', mode = 'markers', sizes = c(5,50),
        marker = list(symbol = 'circle', sizemode = 'diameter',
                      line = list(width = 2, color = '#FFFFFF')),
        text = ~paste('ticker:', symbol, '<br>sector:', sector, '<br>return:', scales::percent(cumulative_ret),
                      '<br>annualized sd:', scales::percent(avg_sd))) %>% 
    layout(xaxis = list(tickformat = "%", title = "annualized sd"),
           yaxis = list(tickformat = "%", title = "cumulative return"),
           plot_bgcolor = "#000000",
         paper_bgcolor = "#000000",
         font = list(color="#ffffff"),
      showlegend = F) %>% 
  animation_opts(1000) %>% 
  animation_button(visible = F) %>% 
  config(displayModeBar = F)
```


```{r}
evolved_holdings <- 
iecs_holdings %>% 
  bind_rows(iede_holdings, iefn_holdings,iehs_holdings, ietc_holdings, ieih_holdings, ieme_holdings) %>% 
  group_by(fund) %>% 
  na.omit() %>% 
  clean_names()
```

### Add sector and industry labels

The evolved funds use NLP to classify tickers differently than the GICS standards but I'm curious how the GICS sectors and industries map back to the funds. I want to add the sector and industry labels to the the joined funds we imported above but first I need a data frame with the tickers, sectors and industries. The `tq_exchange()` function from `tidyquant` allows us to pull in tickers, sectors and industries by exchange.  

```{r}

nyse <- tq_exchange("NYSE")

nyse_sectors <- 
  nyse %>% 
  select(ticker = symbol, sector, industry)

nasdaq <- tq_exchange("NASDAQ")

nasdaq_sectors <- 
  nasdaq %>% 
  select(ticker = symbol, sector, industry)

amex <- tq_exchange("AMEX")

amex_sectors <- 
  amex %>% 
  select(ticker = symbol, sector, industry)

nyse_sectors <- read_rds("nyse_sectors.RDS")
amex_sectors <- read_rds("amex_sectors.RDS")
nasdaq_sectors <- read_rds("nasdaq_sectors.RDS")
```

Now we can use `left_join()` to add the sector/industry labels to the `evolved_holdings` object. We will join them by the common column called `Ticker`. 

```{r}
evolved_holdings %>%
  left_join(nyse_sectors, by = "ticker") %>% 
  left_join(nasdaq_sectors, by = "ticker") %>% 
  #left_join(amex_sectors, by = "ticker") %>% 
  select(contains("sector"), contains("industry"), ticker)
```

Note that a slight issue arose for us - we joined two sets of labels, that means we tried to add two columns called `sector` and two columns called `industry`. `left_join()` notices that and gives the columns unique names. But now we want to combine them to one column called `sector` and one called `industry`. We can use `case_when()` for this.

```{r}
evolved_holdings %>%
  left_join(nyse_sectors, by = "ticker") %>% 
  left_join(nasdaq_sectors, by = "ticker") %>% 
  mutate(sector = case_when(!(is.na(sector.x)) ~ sector.x,
                            !(is.na(sector.y)) ~ sector.y),
         industry = case_when(!(is.na(industry.x)) ~ industry.x,
                            !(is.na(industry.y)) ~ industry.y)) %>% 
  select(-sector.x, -sector.y, -industry.x, -industry.y) %>% 
  select(ticker, contains("sector"), contains("industry"))
```

Are we 100% sure all of our tickers received a label? Let's check.

```{r}
evolved_holdings %>%
  left_join(nyse_sectors, by = "ticker") %>% 
  left_join(nasdaq_sectors, by = "ticker") %>% 
  mutate(sector = case_when(!(is.na(sector.x)) ~ sector.x,
                            !(is.na(sector.y)) ~ sector.y),
         industry = case_when(!(is.na(industry.x)) ~ industry.x,
                            !(is.na(industry.y)) ~ industry.y)) %>% 
  select(-sector.x, -sector.y, -industry.x, -industry.y) %>% 
  select(ticker, contains("sector"), contains("industry")) %>% 
  filter(is.na(sector)) %>% 
  arrange(desc(ticker))
```

Looks like we missed 28 tickers. `XTSLA` is a Blackrock money market fund , `USD` is cash and those are 14 of the tickers we missed. Remember we have 7 funds so it seems that each fund holds some cash and some allocation a money market fund. I couldn't find a better way than to search for the rest of the tickers and manually enter their sectors and industries.


```{r}
evolved_holdings %>%
  left_join(nyse_sectors, by = "ticker") %>% 
  left_join(nasdaq_sectors, by = "ticker") %>% 
  mutate(sector = case_when(!(is.na(sector.x)) ~ sector.x,
                            !(is.na(sector.y)) ~ sector.y,
                            ticker == "USD" ~ "cash",
                            ticker == "XTSLA" ~ "cash",
                            ticker == "CBOE" ~ "Finance",
                            ticker == "IBKR" ~ "Finance",
                            ticker == "NHC" ~ "Health Care",
                            ticker == "SGYPQ" ~ "Health Care",
                            ticker == "TMP" ~ "Finance",
                            ticker == "PRK" ~ "Finance",
                            ticker == "LGFA" ~ "Consumer Services",
                            ticker == "LGFB" ~ "Consumer Services",
                            ticker == "BFB" ~ "Consumer Staples",
                            ticker == "BFA" ~ "Consumer Staples",
                            ticker == "JWA" ~ "Communication Services",
                            ticker == "BRKB" ~ "Financials"),
         industry = case_when(!(is.na(industry.x)) ~ industry.x,
                            !(is.na(industry.y)) ~ industry.y,
                            ticker == "USD" ~ "cash",
                            ticker == "XTSLA" ~ "cash",
                            ticker == "CBOE" ~ "Capital Markets",
                            ticker == "IBKR" ~ "Capital Markets",
                            ticker == "NHC" ~ "Health Care Providers & Services",
                            ticker == "SGYPQ" ~ "Pharmaceuticals, Biotech and Life Sciences",
                            ticker == "TMP" ~ "Banks",
                            ticker == "PRK" ~ "Banks",
                            ticker == "LGFA" ~ "Entertainment",
                            ticker == "LGFB" ~ "Entertainment",
                            ticker == "BFB" ~ "Beverages",
                            ticker == "BFA" ~ "Beverages",
                            ticker == "JWA" ~ "Media",
                            ticker == "BRKB" ~ "Diversified Financial Services")) %>% 
  select(-sector.x, -sector.y, -industry.x, -industry.y) %>% 
  filter(is.na(sector)) 
```

Looks good, let's save the labeled object as `evolved_labeled`.

```{r}
evolved_labeled <-
evolved_holdings %>%
  left_join(nyse_sectors, by = "ticker") %>% 
  left_join(nasdaq_sectors, by = "ticker") %>% 
  mutate(sector = case_when(!(is.na(sector.x)) ~ sector.x,
                            !(is.na(sector.y)) ~ sector.y,
                            ticker == "USD" ~ "cash",
                            ticker == "XTSLA" ~ "cash",
                            ticker == "CBOE" ~ "Finance",
                            ticker == "IBKR" ~ "Finance",
                            ticker == "NHC" ~ "Health Care",
                            ticker == "SGYPQ" ~ "Health Care",
                            ticker == "TMP" ~ "Finance",
                            ticker == "PRK" ~ "Finance",
                            ticker == "LGFA" ~ "Consumer Services",
                            ticker == "LGFB" ~ "Consumer Services",
                            ticker == "BFB" ~ "Consumer Staples",
                            ticker == "BFA" ~ "Consumer Staples",
                            ticker == "JWA" ~ "Communication Services",
                            ticker == "BRKB" ~ "Financials"),
         industry = case_when(!(is.na(industry.x)) ~ industry.x,
                            !(is.na(industry.y)) ~ industry.y,
                            ticker == "USD" ~ "cash",
                            ticker == "XTSLA" ~ "cash",
                            ticker == "CBOE" ~ "Capital Markets",
                            ticker == "IBKR" ~ "Capital Markets",
                            ticker == "NHC" ~ "Health Care Providers & Services",
                            ticker == "SGYPQ" ~ "Pharmaceuticals, Biotech and Life Sciences",
                            ticker == "TMP" ~ "Banks",
                            ticker == "PRK" ~ "Banks",
                            ticker == "LGFA" ~ "Entertainment",
                            ticker == "LGFB" ~ "Entertainment",
                            ticker == "BFB" ~ "Beverages",
                            ticker == "BFA" ~ "Beverages",
                            ticker == "JWA" ~ "Media",
                            ticker == "BRKB" ~ "Diversified Financial Services")) %>% 
  select(-sector.x, -sector.y, -industry.x, -industry.y) 
```

Almost ready to visualize these fund holdings. I have in mind to build a tree that proceeds from fund, to sector, to industry, to asset weights. There's a column for weights but it's a bit too fine grained, basically a continuous variable. Let's organize the fund holdings into buckets based on their weights using the `ntile()` function from `dplyr`. In this case, I want to create 5 buckets and that function will try to put the same number of assets into each bucket, based on their weights. Here's how that gets coded. I'll use `slice(1:5)` to show the top 5 holdings in bin 1 for each fund.

```{r}
evolved_labeled %>% 
  mutate(buckets = ntile(desc(weight_percent), 5),
         buckets_label = str_c("weights_bin_", buckets, sep = "")) %>% 
  select(buckets, buckets_label, fund, everything()) %>% 
  slice(1, 30, 50, n())

evolved_labeled_bucketed <-
evolved_labeled %>% 
  mutate(buckets = ntile(desc(weight_percent), 5),
         buckets = str_c("weights_bin_", buckets, sep = ""))

write_rds(evolved_labeled_bucketed, file = "evolved_labeled_bucketed.RDS")
```


Now let's visualize these funds using the new [collapsibleTree](https://adeelk93.github.io/collapsibleTree/). We set the hierarchy of nodes with the `hierarchy` argument and identify what we wish to display when a user hovers with the `attribute` argument. 

```{r, fig.width=5}
evolved_labeled_bucketed %>% 
    collapsibleTreeSummary(
      root = "evolved",
      hierarchy = c("fund","sector", "industry", "buckets", "ticker"),
      attribute = "shares",
      zoom = FALSE
    )
```




e voila! 