---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(readxl)
library(plotly)
library(riingo)
library(tidyquant)
library(tidyverse)
library(timetk)

library(ggtext)
riingo_set_token("97d26898fe27141ef524c1fa3ab0d3f0c3c59b09")

pins::board_register(board = "rsconnect", server = "https://colorado.rstudio.com/rsc", key = "fAEzqlEWVZM8GEQAkoQiKTOYgMXrAvZ2")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

For today we'll be analyzing the holdings of the Vanguard's total market ETF, VTI. We can grab the full holdings from Vangaurd's website [here](https://advisors.vanguard.com/investments/products/vti/vanguard-total-stock-market-etf#portfolio) and import a CSV of the fund's full holdings. 

Here's a snapshot of the raw CSV. 

![]()

Let's import that into RStudio using `read_csv()`.

```{r}
vti_total_mkt_holdings <- 
  read_csv("Top_holdings_Total_Stock_Market_ETF.csv", 
           col_types = cols(COUNTRY = col_skip(), 
                            `SECURITY DEPOSITORY RECEIPT TYPE` = col_skip(), 
                            SEDOL = col_skip(), X1 = col_skip()), 
           skip = 7) %>% 
  janitor::clean_names()
```

We have 3,559 tickers and our analysis will go back about 3 years today. Let's grab those price histories using the excellent price data source [tiingo](https://www.tiingo.com/). There's an R package called `riingo` that makes it easy to access the `tiingo` API via a function called `riingo_prices()`. We'll supply the first 1000 tickers to that function and pull in 3-years of price history. 

```{r, echo=FALSE}
vti_prices_1_1500 <-  read_rds("vti_prices_1_1000.RDS")
```


```{r, eval=FALSE}
# Set start and end dates for the price import
end <- today() %>% ymd()
start <- end - years(3)

# pull tickers and pass to riingo_prices()
vti_prices_1_1000 <-
  vti_total_mkt_holdings %>%
  mutate(ticker = str_replace(ticker, "BRK.B", "BRK-B")) %>%
  slice(1:1000) %>%
  pull(ticker) %>%
  riingo_prices(start_date = start, end_date = end)

write_rds(vti_prices_1_1000, "vti_prices_1_1000.RDS")

vti_prices_1001_2000 <-
  vti_total_mkt_holdings %>%
  janitor::clean_names() %>%
  # mutate(ticker = str_replace(ticker, "BRK.B", "BRK-B")) %>%
  slice(1001:2000) %>%
  pull(ticker) %>%
  riingo_prices(start_date = start, end_date = end)

vti_prices_1_2000 <- 
vti_prices_1_1000 %>% 
  bind_rows(vti_prices_1001_2000)
```


I have an R object that stores the more traditional exchange sector labels and maps them to several thousand tickers. Here's a peek: 

```{r}

exchange_tickers_sectors <- read_rds("exchange_tickers_sectors.RDS")

exchange_tickers_sectors %>% 
  head()
```

```{r}
vti_prices_1_1000 %>% 
  select(ticker, date, close, volume) %>% 
  left_join(vti_total_mkt_holdings %>% select(ticker, sector), by = c("ticker" = "ticker")) %>% 
  left_join(exchange_tickers_sectors %>% select(ticker, exchange_sector = sector)) %>%  
  group_by(ticker) %>% 
  mutate(sma_50 = roll_mean(as.matrix(close), 50, complete_obs = T),
         sma_200 = roll_mean(as.matrix(close), 200, complete_obs = T),
         sma_50_greater_than_sma_200 = case_when(sma_50 > sma_200 ~ 1, 
                                                 TRUE ~ 0),
         current_greater_than_sma_50 = case_when(close > sma_50 ~ 1, 
                                                 TRUE ~ 0)) %>%
  na.omit() %>% 
  slice(n()) %>% 
  group_by(exchange_sector) %>% 
  count(sma_50_greater_than_sma_200) %>%  
  mutate(percent = n/sum(n), 
         label = case_when(sma_50_greater_than_sma_200 == 1 ~ "sma50 above sma200", 
                           TRUE ~ "sma50 below sma200")) %>% 
  group_by(label) %>% 
  filter(!(str_detect(exchange_sector, 'cash|Cash'))) %>% 
  mutate(exchange_sector = str_wrap(exchange_sector, width = 13)) %>%
  ggplot(aes(x = exchange_sector, y = percent, fill = label)) + 
  geom_col(width = .3, position = position_dodge(width = .5)) +
  scale_y_continuous(labels = scales::percent_format(), breaks = scales::pretty_breaks(n = 10)) +
    labs(x = "", y = "", title = "Percent SMA 50 <span style='color:blue'>above</span> or <span style='color:pink'>below</span> SMA 200", subtitle = str_glue("as of {date_filter}") ) +
  scale_fill_manual(values = c( "blue", "pink")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title = element_markdown(hjust = .5),
        plot.subtitle = element_text(hjust = .5))
```



```{r}

vti_sma50_sma_200_for_gif <- 
vti_prices_1_1000 %>% 
  select(ticker, date, close, volume) %>% 
  left_join(vti_total_mkt_holdings %>% select(ticker, sector), by = c("ticker" = "ticker")) %>% 
  left_join(exchange_tickers_sectors %>% select(ticker, exchange_sector = sector)) %>%  
  group_by(ticker) %>% 
  mutate(sma_50 = roll_mean(as.matrix(close), 50, complete_obs = T),
         sma_200 = roll_mean(as.matrix(close), 200, complete_obs = T),
         sma_50_greater_than_sma_200 = case_when(sma_50 > sma_200 ~ 1, 
                                                 TRUE ~ 0),
         current_greater_than_sma_50 = case_when(close > sma_50 ~ 1, 
                                                 TRUE ~ 0)) %>%
  na.omit() %>% 
  filter(date > "2017-10-04") %>% 
  group_by(exchange_sector, date) %>% 
  count(sma_50_greater_than_sma_200) %>% 
  mutate(percent = n/sum(n), 
         trend = case_when(sma_50_greater_than_sma_200 == 1 ~ "sma50 above sma200", 
                           TRUE ~ "sma50 below sma200"),
         month = month(date)) %>% 
  group_by(trend) %>% 
  filter(!(str_detect(exchange_sector, 'cash|Cash'))) %>% 
  mutate(exchange_sector = str_wrap(exchange_sector, width = 13)) %>%
  group_by(strftime(date, "%Y-%m")) %>% #Groups by the yearmonths
  filter(date == max(date)) %>% 
  ungroup()

write_rds(vti_sma50_sma_200, "vti_sma50_sma_200.RDS")

date_choices <-
  vti_sma50_sma_200 %>% 
  distinct(date) %>% 
  pull(date)

write_rds(date_choices, "date_choices.RDS")
```



```{r}
library(gganimate)
library(gifski)
library(png)

p_1 <-  
  vti_sma50_sma_200 %>% 
 filter(date > "2018-01-01") %>% 
  group_by(exchange_sector, date) %>% 
  count(sma_50_greater_than_sma_200) %>% 
  mutate(percent = n/sum(n), 
         trend = case_when(sma_50_greater_than_sma_200 == 1 ~ "sma50 above sma200", 
                           TRUE ~ "sma50 below sma200"),
         month = month(date)) %>% 
  group_by(trend) %>% 
  filter(!(str_detect(exchange_sector, 'cash|Cash'))) %>% 
  mutate(exchange_sector = str_wrap(exchange_sector, width = 13)) %>%
  group_by(strftime(date, "%Y-%m")) %>% #Groups by the yearmonths
  filter(date == max(date)) %>% 
  ggplot(aes(x = exchange_sector, y = percent, fill = trend)) + 
  geom_col(width = .3, position = position_dodge(width = .5), show.legend = F) +
  scale_y_continuous(labels = scales::percent_format(), breaks = scales::pretty_breaks(n = 10)) +
    #labs(x = "", y = "", title = "Percent SMA 50 above or below SMA 200", subtitle = str_glue("as of {date}") ) +
  scale_fill_manual(values = c( "darkgreen", "pink")) +
  theme_minimal() +
  transition_time(date) +
  labs(title = "Percent SMA 50 <span style='color:darkgreen'>above</span> or <span style='color:pink'>below</span> SMA 200", subtitle = " {format(frame_time, '%Y-%m-%d')}", y= "", x = "") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title = element_markdown(hjust = .5),
        plot.subtitle = element_text(hjust = .5))

  animate(p_1, fps = 5, renderer = gifski_renderer(loop = TRUE))
```

```{r}
p <- 
  vti_sma50_sma_200 %>%  
  ggplot(aes(x = exchange_sector, y = percent, fill = trend)) + 
  geom_col(width = .3, position = position_dodge(width = .5), show.legend = F) +
  scale_y_continuous(labels = scales::percent_format(), breaks = scales::pretty_breaks(n = 10)) +
    #labs(x = "", y = "", title = "Percent SMA 50 above or below SMA 200", subtitle = str_glue("as of {date}") ) +
  scale_fill_manual(values = c( "darkgreen", "pink")) +
  theme_minimal() +
  transition_states(date,
                    transition_length = 5,
                    state_length = 5) +
  labs(title = "Percent SMA 50 <span style='color:darkgreen'>above</span> or <span style='color:pink'>below</span> SMA 200", subtitle = " {closest_state}", y= "", x = "") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title = element_markdown(hjust = .5),
        plot.subtitle = element_text(hjust = .5)) +
  ease_aes('cubic-in-out')  +
  enter_fade() +
  exit_shrink()

animate(p, fps = 5, renderer = gifski_renderer(loop = TRUE))
```


```{r}

  
sma50_sma200_fun <- function(price_tibble, as_of_date){
date_filter <- as_of_date
(
price_tibble %>% 
  filter(date == date_filter) %>% 
  group_by(exchange_sector) %>% 
  count(sma_50_greater_than_sma_200) %>%  
  mutate(percent = n/sum(n), 
         trend = case_when(sma_50_greater_than_sma_200 == 1 ~ "sma50 above sma200", 
                           TRUE ~ "sma50 below sma200"),
         tooltip_percent = scales::percent(percent),
         month_date = strftime(date, "%Y-%m")) %>% 
  group_by(trend) %>% 
  filter(!(str_detect(exchange_sector, 'cash|Cash'))) %>% 
  mutate(exchange_sector = str_wrap(exchange_sector, width = 13)) %>%
  ggplot(aes(x = exchange_sector, y = percent, fill = trend, text = paste("sector:", exchange_sector, "<br>trend:", trend, "<br>percent", tooltip_percent))) + 
  geom_col(width = .3, position = position_dodge(width = .5)) +
  scale_y_continuous(labels = scales::percent_format(), breaks = scales::pretty_breaks(n = 10)) +
    labs(x = "", y = "", title = "Percent SMA 50 above or below SMA 200", subtitle = str_glue("as of {date_filter}") ) +
  scale_fill_manual(values = c( "lightgreen", "pink")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(hjust = .5))
) %>% 
  ggplotly(tooltip = "text")
}

```

JKR extensions from here: 

Rolling percents over time

```{r}
vti_prices_1_1000 %>% 
  select(ticker, date, close, volume) %>% 
  left_join(vti_total_mkt_holdings %>% select(ticker, sector), by = c("ticker" = "ticker")) %>% 
  left_join(exchange_tickers_sectors %>% select(ticker, exchange_sector = sector)) %>%  
  group_by(ticker) %>% 
  mutate(sma_50 = roll_mean(as.matrix(close), 50, complete_obs = T),
         sma_200 = roll_mean(as.matrix(close), 200, complete_obs = T),
         sma_50_greater_than_sma_200 = case_when(sma_50 > sma_200 ~ 1, 
                                                 TRUE ~ 0),
         current_greater_than_sma_50 = case_when(close > sma_50 ~ 1, 
                                                 TRUE ~ 0)) %>%
  na.omit() %>% 
  group_by(exchange_sector, date) %>% 
  count(sma_50_greater_than_sma_200) %>%  
  mutate(percent = n/sum(n), 
         trend = case_when(sma_50_greater_than_sma_200 == 1 ~ "sma50 above sma200", 
                           TRUE ~ "sma50 below sma200")) %>% 
  group_by(trend) %>% 
  filter(!(str_detect(exchange_sector, 'cash|Cash'))) %>% 
  mutate(exchange_sector = str_wrap(exchange_sector, width = 13)) %>% 
  filter(trend == "sma50 above sma200") %>% 
  ggplot(aes(x = date, y = percent, color = exchange_sector)) +
  geom_line(show.legend = F) +
  geom_hline(yintercept = .5) +
  labs(y = "percent sma50 above sma200") +
  scale_y_continuous(labels = scales::percent, breaks = scales::pretty_breaks(n = 10)) +
  facet_wrap(~exchange_sector)
```

```{r}
(
vti_prices_1_1000 %>% 
  select(ticker, date, close, volume) %>% 
  left_join(vti_total_mkt_holdings %>% select(ticker, sector), by = c("ticker" = "ticker")) %>% 
  left_join(exchange_tickers_sectors %>% select(ticker, exchange_sector = sector)) %>%  
  group_by(ticker) %>% 
  mutate(sma_50 = roll_mean(as.matrix(close), 50, complete_obs = T),
         sma_200 = roll_mean(as.matrix(close), 200, complete_obs = T),
         sma_50_greater_than_sma_200 = case_when(sma_50 > sma_200 ~ 1, 
                                                 TRUE ~ 0),
         current_greater_than_sma_50 = case_when(close > sma_50 ~ 1, 
                                                 TRUE ~ 0)) %>%
  na.omit() %>% 
  group_by(exchange_sector, date) %>% 
  count(sma_50_greater_than_sma_200) %>%  
  mutate(percent = n/sum(n), 
         trend = case_when(sma_50_greater_than_sma_200 == 1 ~ "sma50 above sma200", 
                           TRUE ~ "sma50 below sma200")) %>% 
  group_by(trend) %>% 
  filter(!(str_detect(exchange_sector, 'cash|Cash'))) %>% 
  mutate(exchange_sector = str_wrap(exchange_sector, width = 13),
         date = ymd(date)) %>% 
  filter(trend == "sma50 above sma200") %>% 
  ggplot(aes(x = date, y = percent, color = exchange_sector)) +
  geom_line(show.legend = F) +
  geom_hline(yintercept = .5) +
  labs(title = "percent sma 50 above sma 200", y = "", x = "") +
  scale_y_continuous(labels = scales::percent, breaks = scales::pretty_breaks(n = 10)) +
  scale_x_date(breaks = scales::pretty_breaks(n = 10)) +
  theme(axis.text.x = element_text(angle = 60))
) %>% 
  ggplotly()
```

choose different metrics, like current price above SMA or rolling volumes. 
choose your date where you'd like to see this stat
rolling over time by sector, where the line is percentage above, and maybe negative line for percentage below? or all lines one chart by color for sector
over lay future returns - is this giving us a signal for the index? can we find a sector rotation or trading strategy here? like, top three sector percentages, rotate towards those.

```{r}
ivv_holdings <- 
read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund", 
    col_types = cols(ISIN = col_skip(), SEDOL = col_skip()), 
    skip = 10) %>% 
  mutate(fund = "ivv", 
         download_date = ymd(Sys.Date()),
         Ticker = str_replace(Ticker, "BRKB", "BRK-B"))

# ivv_prices <- 
# ivv_holdings %>% 
#   slice(1:500) %>% 
#   pull(Ticker) %>% 
#   tq_get(get = "stock.prices", from = "2019-01-01") %>% 
#   left_join(ivv_holdings %>% select(Ticker, Sector), by = c("symbol" = "Ticker"))


#write_rds(ivv_prices, "ivv_prices.RDS")

ivv_prices %>% 
  distinct(symbol)

ivv_prices %>% 
  group_by(symbol) %>% 
  mutate(sma_50 = roll_mean(as.matrix(close), 50, complete_obs = T),
         sma_200 = roll_mean(as.matrix(close), 200, complete_obs = T),
         short_greater_than_long = case_when(sma_50 > sma_200 ~ 1, 
                                   TRUE ~ 0),
         current_greater_than_short = case_when(close > sma_50 ~ 1, 
                                   TRUE ~ 0)) %>% 
  slice(n()) %>% 
  select(symbol, Sector, short_greater_than_long) %>% 
  group_by(Sector) %>% 
  count(short_greater_than_long) %>% 
  #ungroup() %>% 
  mutate(percent = n/sum(n), 
         label = case_when(short_greater_than_long == 0 ~ "sma50 below sma200", 
                           TRUE ~ "sma50 above sma200")) %>% 
  group_by(label) %>% 
  # mutate(Sector = case_when(Sector == "Information Technology" ~ "Info Tech",
  #                           Sector == "Consumer Discretionary" ~ "Cons Disc",
  #                           Sector == "Communication" ~ "Communic",
  #                           Sector == "Consumer Staples" ~ "Cons Staples",
  #                           TRUE ~ Sector)) %>%
  filter(Sector != "Cash and/or Derivatives") %>% 
  mutate(Sector = str_wrap(Sector, width = 10)) %>%
  ggplot(aes(x = Sector, y = percent, fill = label)) + 
  geom_col(width = .3, position = position_dodge(width = .5)) +
  scale_y_continuous(labels = scales::percent_format(), breaks = scales::pretty_breaks(n = 10)) +
    labs(x = "", y = "", title = "Percent SMA 50 above or below SMA 200", subtitle = "by sector" ) +
  scale_fill_manual(values = c( "blue", "pink")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(hjust = .5))
```

