---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(plotly)
library(riingo)
library(roll)
library(tidyquant)
library(tidyverse)
library(timetk)

riingo_set_token("97d26898fe27141ef524c1fa3ab0d3f0c3c59b09")

pins::board_register(board = "rsconnect", server = "https://colorado.rstudio.com/rsc", key = "fAEzqlEWVZM8GEQAkoQiKTOYgMXrAvZ2")

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


# January holdings 2013 - 2019

```{r}

spy_holding_jan_2013 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20130131", 
    skip = 10) %>% 
  mutate(date = ymd("2013-01-31"),
         Ticker = str_remove(Ticker, "\\*"))  %>% 
  select(date, ticker = Ticker, weight = `Weight (%)`, sector = Sector, price = Price) 

spy_holding_jan_2014 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20140131", 
    skip = 10) %>% 
  mutate(date = ymd("2014-01-31"),
         Ticker = str_remove(Ticker, "\\*")) %>% 
  select(date, ticker = Ticker, weight = `Weight (%)`, sector = Sector, price = Price) 

spy_holding_jan_2015 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20150130", 
    skip = 10) %>% 
  mutate(date = ymd("2015-01-30")) %>% 
  select(date, ticker = Ticker, weight = `Weight (%)`, sector = Sector, price = Price) 

spy_holding_jan_2016 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20160129", 
    skip = 10) %>% 
  mutate(date = ymd("2016-01-29")) %>% 
  select(date, ticker = Ticker, weight = `Weight (%)`, sector = Sector, price = Price) 

spy_holding_jan_2017 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20170131", 
    skip = 10) %>% 
  mutate(date = ymd("2017-01-31")) %>% 
  select(date, ticker = Ticker, weight = `Weight (%)`, sector = Sector, price = Price) 

spy_holding_Jan_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180131", 
    skip = 10) %>% 
  mutate(date = ymd("2018-01-31")) %>% 
  select(date, ticker = Ticker, weight = `Weight (%)`, sector = Sector, price = Price)

spy_holding_jan_2019 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20190131", 
    skip = 10) %>% 
  mutate(date = ymd("2019-01-31")) %>% 
  select(date, ticker = Ticker, weight = `Weight (%)`, sector = Sector, price = Price)


january_holdings <-
spy_holding_jan_2013 %>% 
bind_rows(spy_holding_jan_2014) %>% 
bind_rows(spy_holding_jan_2015) %>% 
bind_rows(spy_holding_jan_2016) %>% 
bind_rows(spy_holding_jan_2017) %>% 
bind_rows(spy_holding_Jan_2018) %>% 
bind_rows(spy_holding_jan_2019) %>% 
  group_by(date) %>% 
  mutate(year = year(date)) 

```

# Monthly 2018 holdings and weights

```{r}
# spy_holding_Jan_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180131", 
#     skip = 10) %>% 
#   mutate(date = ymd("2018-01-31")) %>% 
#   select(date, ticker = Ticker, weight = `Weight (%)`, sector = Sector, price = Price) 

spy_holding_feb_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180228", 
    skip = 10)

spy_holding_mar_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180329", 
    skip = 10)

spy_holding_apr_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180430", 
    skip = 10)

spy_holding_may_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180531", 
    skip = 10)

spy_holding_jun_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180629", 
    skip = 10)

spy_holding_jul_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180731", 
    skip = 10)

spy_holding_aug_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180831", 
    skip = 10)

spy_holding_sep_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20180928", 
    skip = 10)

spy_holding_oct_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20181031", 
    skip = 10)

spy_holding_nov_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20181130", 
    skip = 10)

spy_holding_dec_2018 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20181231", 
    skip = 10)
```

# Some random dates

```{r}
spy_holding_dec_2017 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20171229", 
    skip = 10) %>% 
  select(date, ticker = Ticker, weight = `Weight (%)`, sector = Sector, price = Price) 


spy_holding_july_2019 <- read_csv("https://www.ishares.com/us/products/239726/ishares-core-sp-500-etf/1467271812596.ajax?fileType=csv&fileName=IVV_holdings&dataType=fund&asOfDate=20190731", 
    skip = 10)


```

## Save fund holdings to list

```{r}

# spy_holdings_list <-
#   list(spy_holding_jan_2013,
#        spy_holding_jan_2014,
#        spy_holding_jan_2015,
#        spy_holding_jan_2016,
#        spy_holding_jan_2017,
#        spy_holding_Jan_2018,
#        spy_holding_jan_2019)


library(pins)
# 
pin(ivv_holdings_list, description = "List of tibbles with holdings for sp500 2013 through 2019", board = "rsconnect")


ivv_holdings_list  <- pins::pin_get("jonathan/ivv-holdings-list", board = "rsconnect")

years <- seq(2013, 2019, by = 1)

names(ivv_holdings_list) <- years

ivv_holdings_list <- 
 spy_holdings_list %>% 
  map(~mutate(.x, ticker = str_replace(ticker, "BRKB", "BRK-B")))

ivv_holding_jan_2018 <- 
  ivv_holdings_list %>% 
  .[['2018']]  #%>%
  # mutate(ticker = str_replace(ticker, "BRKB", "BRK-B")) %>%
  # filter(ticker %in% c("AET", "MON"))
```

## Get constituent returns for one year

First, we import the daily prices for each constituent for 2018 and then use `mutate()` to calculate daily returns. 

```{r}
ivv_prices_tiingo <- 
ivv_2018_tickers %>% 
  riingo_prices(start_date = "2018-01-01", end_date = "2018-12-31", resample_frequency = "monthly")
  select(ticker = symbol, date, adjusted) %>% 
  group_by(ticker) %>% 
  mutate(daily_return = adjusted/lag(adjusted, 1) - 1) %>% 
  summarise(dailyly_sd = sd(daily_return, na.rm = T),
            yearly_return  = last(adjusted)/first(adjusted) -1) %>% 
  left_join(ivv_holding_jan_2018)
```

We can create a function that takes as an input the year, the number of holdings a tibble of holdings and weights. The function then imports prices for that year and calculates daily returns. Then it calculates standard deviation of returns for each ticker and the yearly return for each ticker. This allows us to visualize on a year by year basis, the sd and return for the top N holdings in the fund or portfolio.


Create the function:

```{r}

get_yearly_returns_sd <- function(holdings_tibble, year = 2013, top_n_holdings = 10){
  from_date = str_glue(year, "-01-01", sep = "")
  to_date = str_glue(year, "-12-31", sep = "")
  
  holdings_tibble %>% 
  slice(1:top_n_holdings) %>% 
  pull(ticker) %>% 
  tq_get(from = from_date, to = to_date) %>% 
  select(ticker = symbol, date, adjusted) %>% 
  group_by(ticker) %>% 
  mutate(daily_return = adjusted/lag(adjusted, 1) - 1) %>% 
  summarise(yearly_sd = sd(daily_return, na.rm = T),
            yearly_return  = last(adjusted)/first(adjusted) -1) %>% 
  left_join(holdings_tibble)
}
```

Test the function on one year of holdings.


```{r}

spy_holding_jan_2019 <- 
  spy_holdings_list %>% 
  .[["2019"]]  

get_yearly_returns_sd(spy_holding_jan_2019, 2019, 10)
```

Create a vector of years that we can map over and efficiently grab prices, returns summary stats over multiple years.

```{r}

# years <- seq(2013, 2019, by = 1)

# fund_weights_returns <-
#   spy_holdings_list %>% 
#   map_dfr(get_yearly_returns_sd, 1) %>% 
#   group_by(date) %>% 
#   mutate(year = year(date),
#          weight = weight/100)

# get_yearly_returns_sd(spy_holding_jan_2019, top_n_holdings = 5, year = years[7])

fund_weights_returns_top_10 <-
spy_holdings_list %>% 
map2_dfr(years, get_yearly_returns_sd, top_n_holdings = 10) %>% 
  group_by(date) %>% 
  mutate(year = year(date),
         weight = weight/100)

fund_weights_returns_top_40 <-
spy_holdings_list %>% 
map2_dfr(years, get_yearly_returns_sd, top_n_holdings = 40) %>% 
  group_by(date) %>% 
  mutate(year = year(date),
         weight = weight/100)

#pin(fund_weights_returns_top_40, description = "top 40 holdings in sp500, 2013 through 2019, weights, yearly return and sd", board = "rsconnect")

fund_weights_returns_top_40  <- pins::pin_get("jonathan/fund-weights-returns-top-40", board = "rsconnect")
```


# Visualize annual risk and returns over time

```{r}
colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')

ax <- list(
  title = "",
  zeroline = FALSE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)


fund_weights_returns_top_40 %>% 
  plot_ly(x = ~yearly_sd, y = ~yearly_return, color = ~ticker, size = ~weight, frame = ~date,
        type = 'scatter', 
        mode = 'markers', 
        sizes = c(5,50),
        marker = list(symbol = 'circle', 
                      sizemode = 'diameter',
                      opacity = 0.5,
                      line = list(width = 2, 
                                  color = '#FFFFFF')),
        text = ~paste('ticker:', ticker, '<br>sector:', sector, 
                      '<br>weight:', scales::percent(weight), 
                      '<br>return:', scales::percent(yearly_return),
                      '<br>risk:', scales::percent(yearly_sd)),
        hoverinfo = 'text') %>%
 # add_markers(color = ~sector, size = ~weight)
    layout(xaxis = list(tickformat = "%", title = "risk"),
           yaxis = list(tickformat = "%", title = "return"),
           plot_bgcolor = "#000000",
         paper_bgcolor = "#000000",
         font = list(color="#ffffff"),
      showlegend = T) %>% 
  animation_opts(1500) %>% 
  animation_button(visible = F) %>% 
  config(displayModeBar = F)
```


```{r}
library(ggrepel)
year_text_repel <- 2016
top_n_text_repel <- 20

fund_weights_returns_top_40 %>% 
  filter(year == year_text_repel) %>% 
  arrange(desc(weight)) %>% 
  ggplot(aes(x = yearly_sd, y = yearly_return, color = sector)) + 
  geom_point(
    data = function(x) top_n(x, top_n_text_repel, wt = weight),
    aes(size = weight), 
    alpha = 1
  ) +
  geom_text_repel(
    data = function(x) top_n(x, top_n_text_repel, wt = weight), #filter(x, row_number() <= 10L),
    aes(label = ticker),
    point.padding = .2,
    box.padding = .3,
    force = 1,
    min.segment.length = 0
  ) +
  geom_point(
    data = function(x) top_n(x, -top_n_text_repel * 2, wt = weight),
    aes(size = weight), 
    alpha = .4
  ) +
  theme_minimal(14)  +
  labs(
    x = "risk",
    y = "return",
    title = paste("ishare Core SP 500 Top", top_n_text_repel, "holdings", year_text_repel, sep = " ")
  ) +
  theme(
    legend.position = "none",
    # panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
```


```{r}

ivv_2018_tickers <- 
ivv_holding_jan_2018 %>% 
  select(ticker) %>% 
  distinct() %>% 
  pull()


ivv_prices_tiingo <- 
ivv_2018_tickers %>% 
  riingo_prices(start_date = "2018-01-01", end_date = "2018-12-31", resample_frequency = "monthly")

ivv_tickers_that_worked <- 
ivv_prices_tiingo %>% 
  distinct(ticker) %>% 
  pull()

problem_tickers <- 
setdiff(ivv_2018_tickers, ivv_tickers_that_worked)

problem_weight <-
ivv_holding_jan_2018 %>% 
  filter(ticker %in% problem_tickers) %>% 
  summarise(sum(weight, na.rm = T)/100) %>% 
  pull()
```


# Calc daily returns, spread wider, to xts

```{r}

ivv_2018_component_returns <-
ivv_prices_tiingo %>% 
  select(ticker, date, adjClose) %>% 
  group_by(ticker) %>% 
  mutate(daily_return = adjClose/lag(adjClose, 1) - 1) %>% 
  select(date, ticker, daily_return) %>% 
  pivot_wider(names_from = "ticker", values_from = "daily_return") %>% 
  tk_xts(date_var = date)
```

# Extract weights

```{r}
  
ivv_2018_weights <- 
  ivv_holding_jan_2018 %>% 
  filter(ticker %in% ivv_tickers_that_worked) %>%
  mutate(weight = (weight/100)) %>% 
  pull(weight)

# check what they sum to
1 - sum(ivv_2018_weights, na.rm = T)
```

We are off .0058 because of the problematic tickers. The goal of today's exercise is to understand contributions to portfolio return spread amongst 500 securities. We can live with .0058 of weighting error spread across those 500.


```{r}
ivv_contribution_obj <- 
 Return.portfolio(ivv_2018_component_returns, weights = ivv_2018_weights, rebalance_on = "quarters",  verbose = TRUE)
```


```{r}

to.period.contributions(ivv_contribution_obj$contribution, period = "years") %>% 
  tk_tbl(preserve_index = T, rename_index = "date") %>% 
  pivot_longer(-date, names_to = "ticker", values_to = "contribution") %>% 
  left_join(ivv_holding_jan_2018%>% select(ticker, sector), by = c("ticker" =  "ticker")) %>% 
  arrange(desc(contribution))
```





 